### 2015-09-11 paper1/paper.tex
Sourcefiles for the first 2 papers

\documentclass{elsarticle}
\usepackage{amsmath,amsthm}
\usepackage{bbm}
%\usepackage[all,cmtip]{xy}
\newtheorem{df}{Definition}[section]
\newtheorem{prop}[df]{Proposition}
\newtheorem{lem}[df]{Lemma}
\newtheorem{thm}[df]{Theorem}
\newtheorem{cor}[df]{Corollary}
%\def\N{\mathbbm N}
\def\id{\textrm{id}}
\def\Z{\mathbbm Z}
\def\C{\mathbbm C}
\journal{Topology and its Applications}
\begin{document}
\begin{frontmatter}
\title{Quotients of the order complex $\Delta(\overline{\Pi}_n)$ by subgroups of the Young subgroup $S_1\times S_{n-1}$}
\author[ruelle]{Ralf Donau}
\address[ruelle]{Fachbereich Mathematik, Universit\"at Bremen, 28359 Bremen, Germany}
\ead{ruelle@math.uni-bremen.de}
\begin{abstract}
In this paper we prove that the space $\Delta(\overline{\Pi}_n)/G$ is homotopy equivalent to a wedge of spheres of dimension $n-3$ for all natural numbers $n\geq 3$ and all subgroups $G\subset S_1\times S_{n-1}$. Our argument makes use of Discrete Morse Theory. Furthermore we find a simple formula to compute the number of spheres. For example $\Delta(\overline{\Pi}_n)/S_1\times S_{n-1}$ is homotopy equivalent to a sphere.
\end{abstract}
\begin{keyword}
Discrete Morse Theory\sep regular trisp\sep partial matching\sep acyclic matching\sep Young subgroup
\end{keyword}
\end{frontmatter}
\section{Introduction}
Let $n\geq 3$ and let $\Pi_n$ denote the poset consisting of all partitions of $[n]:=\{1,\dots,n\}$ ordered by refinement, such that the finer partition is the smaller partition. Let $\overline{\Pi}_n$ denote the poset obtained from $\Pi_n$ by removing both the smallest and greatest element, which are $\{\{1\},\dots,\{n\}\}$ and $\{[n]\}$, respectively. Now we define $\Delta(\overline{\Pi}_n)$ to the nerve of the acyclic category $\overline{\Pi}_n$, which is a regular trisp, see \cite[Chapter 10]{buch}.

It is well-known that $\Delta(\overline{\Pi}_n)$ is homotopy equivalent to a wedge of spheres of dimension $n-3$. This result has applications in arrangement theory, since the intersection lattice of a linear braid arrangement in isomorphic to a partition lattice $\Pi_n$, see \cite{braid1}, \cite{braid2}. The following theorem is the first result concerning the topology of the quotient $\Delta(\overline{\Pi}_n)/G$, where $G$ is a non-trivial subgroup of $S_n$. This and related facts have implications in singularity theory, see \cite{sing1}, \cite{sing2}, \cite{sing3}.
\begin{thm}[Kozlov,\cite{clmap}] % p. 198
For any $n\geq 3$, the space $\Delta(\overline{\Pi}_n)/S_n$ is contractible.
\end{thm}
This leads to a general question of determining the homotopy type of $\Delta(\overline{\Pi}_n)/G$ for an arbitrary subgroup $G\subset S_n$. The subgroups which would be especially interesting to consider are the Young subgroups, written as $S_{k_1}\times S_{k_2}\times\dots\times S_{k_r}$ with $k_1+k_2+\dots+k_r=n$.

In this paper we give an answer for all Young subgroups that fix the element $1$, which are of the form $S_1\times S_{k_1}\times\dots\times S_{k_r}$, as a corollary of the more general Proposition \ref{rdthm}. An instance, which we have not yet solved completely, are Young subgroups of the form $S_k\times S_{n-k}$. However, we have computed the first homology with integer coefficients, obtaining the result $H_1(\Delta(\overline{\Pi}_n)/S_k\times S_{n-k};\Z)\cong 0$ for $n\geq 5$.
\section{Discrete Morse Theory}
The definitions of regular trisps, partial matchings, acyclic matchings and foundations of Discrete Morse Theory can be found in \cite{forman}, \cite{clmap} and \cite{buch}. The following two theorems of Discrete Morse Theory are frequently used in our proofs. We also make use of the results in \cite{clmap}.
\begin{thm}[Patchwork Theorem]
\label{patchwork}
Let $\varphi:P\longrightarrow Q$ be an order-preserving map and assume we have acyclic matchings on the subposets $\varphi^{-1}(q)$ for all $q\in Q$. Then the union of these matchings is an acyclic matching on $P$.
\end{thm}
\begin{thm}%[Main Theorem of Discrete Morse Theory for CW Complexes]
\label{morsemain}
Let $\Delta$ be a regular trisp and let $M$ be an acyclic matching on the poset ${\cal F}(\Delta)\setminus\{\hat{0}\}$. Let $c_i$ denote the number of critical $i$-dimensional simplices of $\Delta$. Then:
\begin{enumerate}
\item If the critical simplices form a subcomplex $\Delta_c$ of $\Delta$, then $\Delta$ collapses onto $\Delta_c$.
\item $\Delta$ is homotopy equivalent to a CW complex with $c_i$ cells of dimension $i$.
\end{enumerate}
\end{thm}
The proofs of Theorem \ref{patchwork} and \ref{morsemain} as well as further facts on Discrete Morse Theory can be found in \cite[Chapter 11]{buch}.
\section{The collapsible subtrisp}
\label{allgemein}
Let $n\geq 3$ be a fixed natural number throughout this section. The symmetric group $S_n$ operates on $\Delta(\overline{\Pi}_n)$ in a natural way. In this paper we consider subgroups of the Young subgroup
\[S_1\times S_{n-1}:=\{\sigma\in S_n\mid\sigma(1)=1\}\]
We now consider the regular trisp $\Delta:=\Delta(\overline{\Pi}_n)/G$, where $G\subset S_1\times S_{n-1}$ is a subgroup. Let $\Delta^*$ be that subtrisp of $\Delta$ that we obtain by removing all vertices that are represented by some partition where the block containing $1$ has exactly two elements and any other block is singleton. This condition is independent of the choice of the representative. Such a representative can be written as
\[
v_k:=\{\{1,k\},\{2\},\dots,\widehat{\{k\}},\dots,\{n\}\}
\]
with $k\in\{2,\ldots, n\}$. The element with the hat above is omitted.

Let $R$ be the set of all vertices of $\Delta^*$ where the block containing $1$ is singleton.
\begin{lem}
\label{simplclosure}
There exists an acyclic matching on the poset ${\cal F}(\Delta^*)$ such that a simplex $\sigma\in{\cal F}(\Delta^*)$ is critical if and only if  $V(\sigma)\subset R$.
\end{lem}
\begin{proof}
Let $B$ be the set of all vertices that are not contained in $R$. Now we define a closure map $\varphi:B\longrightarrow R$. Let $b\in B$ and let $\tilde b\in\overline{\Pi}_n$ be a representative of $b$. We modify the blocks in $\tilde b$ as follows. We replace the block $A$ containing $1$ by the two blocks $\{1\}$ and $A\setminus\{1\}$ and take the class of this new partition as $\varphi(b)$. By applying Corollary 2.5 in \cite{clmap} and Proposition 7 in \cite{jule} it is easy to see that this map is well-defined and a closure map\footnote{see Definition 2.1 in \cite{clmap}}. By Theorem 2.2 in \cite{clmap} we get an acyclic matching on ${\cal F}(\Delta^*)$ where a simplex $\sigma$ is critical if and only if  $V(\sigma)\subset R$.
\end{proof}
\begin{lem}
\label{mitte}
There exists an acyclic matching on ${\cal F}(\Delta^*)$ that has only one critical simplex which has dimension $0$. In particular $\Delta^*$ is collapsible.
\end{lem}
\begin{proof}
We define an order-preserving map
\begin{eqnarray*}
\varphi:{\cal F}(\Delta^*)&\longrightarrow&[0,1]\\
\sigma&\longmapsto&
\begin{cases}
0&\text{if $V(\sigma)\subset R$}\\
1&\text{else}
\end{cases}
\end{eqnarray*}
By Lemma \ref{simplclosure} we already have an acyclic matching on $\varphi^{-1}(1)$ with no critical simplices. Now we define an acyclic matching on $\varphi^{-1}(0)$. Let $\alpha$ be the vertex that is represented by the partition $\{\{1\},\{2,\dots,n\}\}$. Let $\sigma$ be a simplex of $\varphi^{-1}(0)$ with $V(\sigma)\not=\{\alpha\}$. It is clear that $\alpha$ can be either deleted from $\sigma$ or uniquely inserted into $\sigma$. The matching rule therefore is the following: add $\alpha$ to a simplex if it is not already there, otherwise remove it. This matching is acyclic, since this rule can be formulated by a closure map that maps each vertex that does not equal $\alpha$ to $\alpha$. We apply Theorem \ref{patchwork} and obtain an acyclic matching on ${\cal F}(\Delta^*)$ that has only one critical simplex; the simplex that has only the vertex $\alpha$. Theorem \ref{morsemain} tells us that $\Delta^*$ is collapsible.
\end{proof}
What about the remaining simplices that are represented by a chain that contains a $v_k$ with $k\in\{2,\ldots, n\}$? Let $V_G$ be the set of all these special vertices. We define a Poset $P_G:=V_G\cup\{0\}$ such that $0$ is the smallest element of $P_G$ and the only element that is comparable with some other element. That means $x,y\in P_G$, $x<y$ implies $x=0$. We define an order-preserving map $\varphi_G:{\cal F}(\Delta)\longrightarrow P_G$ as follows. Let $\sigma\in{\cal F}(\Delta)$, then we map $\sigma$ to $0$ if $\sigma\in{\cal F}(\Delta^*)$. Otherwise we map $\sigma$ to the special vertex of $V_G$ that belongs to $\sigma$, which is unique. By Lemma \ref{mitte} we already have an acyclic matching on $\varphi_G^{-1}(0)$ with one critical cell which has dimension $0$. The next step would be to find acyclic matchings on $\varphi_G^{-1}(v)$ for all $v\in V_G$ and apply Theorem \ref{patchwork}.
\section{The main result}
Let $n\geq 4$, for each $2\leq k\leq n$ we define the embedding $i_k:S_1\times S_{n-2}\longrightarrow S_1\times S_{n-1}$, where
\begin{eqnarray*}
i_k(\sigma):[n]&\longrightarrow&[n]\\
x&\longmapsto&
\begin{cases}
k&\text{if $x=k$}\\
f_k\circ\sigma\circ f_k^{-1}(x)&\text{else}\\
\end{cases}
\end{eqnarray*}
and $f_k$ is the order-preserving bijection $f_k:[n-1]\longrightarrow[n]\setminus\{k\}$. Notice that $\sigma(1)=1$ and hence $i_k(\sigma)(1)=1$. The image of $i_k$ consists of all permutations $\sigma\in S_1\times S_{n-1}$ with $\sigma(k)=k$. It also easy to see that $i_k$ is injective.
\begin{lem}
\label{indlemma}
Let $n\geq 4$ and $G\subset S_1\times S_{n-1}$ be an arbitrary subgroup. Let $2\leq k\leq n$ and assume we have an acyclic matching on ${\cal F}(\Delta(\overline{\Pi}_{n-1})/i_k^{-1}(G))$, where the set of critical simplices consists of one critical simplex of dimension $0$ and $l$ critical simplices of dimension $d$, where $l,d\geq 0$. Let $v\in V_G$ be the vertex that is represented by $v_k$. Then there exists an acyclic matching on $\varphi_G^{-1}(v)$ such that the set of critical simplices consists of $l$ critical simplices of dimension $d+1$.
\end{lem}
\begin{proof}
We define a map
\[
\psi:{\cal F}(\Delta(\overline{\Pi}_{n-1})/i_k^{-1}(G))\longrightarrow\varphi_G^{-1}(v)\setminus\{v\}
\]
as follows. Let $\sigma\in{\cal F}(\Delta(\overline{\Pi}_{n-1})/i_k^{-1}(G))$ be a simplex, then we choose a representative $\tilde\sigma\in{\cal F}(\Delta(\overline{\Pi}_{n-1}))$ and describe what happens to the vertices of $\tilde\sigma$, which are partitions of the set $[n-1]$. Let $p=\{B_1,B_2,\dots,B_r\}$, where $B_1$ is the block that contains $1$, be a partition in the chain $\tilde\sigma$. For $p$ we define $p^*:=\{f_k(B_1)\cup\{k\},f_k(B_2),\dots,f_k(B_r)\}$, which is a partition of set set $[n]$. $\tilde\sigma$ can be written as
\[
\tilde\sigma=(p_1<\ldots<p_t)
\]
We define $\psi(\sigma)$ as the simplex that is represented by the chain $(v_k<p^*_1<\ldots<p^*_t)$.

This map is well-defined, for: Let $(p_1<\ldots<p_t)$ and $(p'_1<\ldots<p'_t)$ be two representatives. That means there exists $g\in i_k^{-1}(G)$ with $p'_i=gp_i$ for all $i\in[t]$. Assume $i\in[t]$, $p_i$ can be written as $p_i=\{B_1,B_2,\dots,B_r\}$ with $1\in B_1$, then we also have $p'_i=\{g(B_1),g(B_2),\dots,g(B_r)\}$ with $1\in g(B_1)$. For $1<j\leq r$ we have $f_k(g(B_j))=i_k(g)(f_k(B_j))$. It also follows $f_k(g(B_1))\cup\{k\}=i_k(g)(f_k(B_1)\cup\{k\})$, since $i_k(g)(k)=k$. Therefore $i_k(g)$ gives us our relation. It is easy to see that $\psi$ is order-preserving.

$\psi$ has an inverse
\[
\psi^{-1}:\varphi_G^{-1}(v)\setminus\{v\}\longrightarrow{\cal F}(\Delta(\overline{\Pi}_{n-1})/i_k^{-1}(G))
\]
that maps a $\sigma\in\varphi_G^{-1}(v)\setminus\{v\}$ as follows. The smallest vertex of $\sigma$ is $v$, we choose a representative $\tilde\sigma=(v_k<p_1<\ldots<p_t)\in{\cal F}(\Delta(\overline{\Pi}_n))$ such that $v_k$ is the smallest partition in the chain $\tilde\sigma$. Let $p=\{B_1,B_2,\dots,B_r\}$, where $B_1$ is the block that contains $1$, be a partition in $\tilde\sigma$. In particular we have $k\in B_1$, since $v_k$ refines any partition in $\tilde\sigma$. For $p$ we define the partition $p^*:=\{f_k^{-1}(B_1\setminus\{k\}),f_k^{-1}(B_2),\dots,f_k^{-1}(B_r)\}$. We define $\psi^{-1}(\sigma)$ as the simplex that is represented by $(p^*_1<\ldots<p^*_t)$.

$\psi^{-1}$ is well-defined, for: Let $(v_k<p_1<\ldots<p_t)$ and $(v_k<p'_1<\ldots<p'_t)$ be two representatives. That means there exists $g\in G$ with $p'_i=gp_i$ for all $i\in[t]$. Clearly we have $g(1)=1$ and $g(k)=k$, and therefore $g$ lies in the image of $i_k$. $i_k^{-1}(g)$ gives us the relation we are looking for. It is easy to see that $\psi^{-1}$ is order-preserving and the inverse of $\psi$.

Via $\psi$ we get an acyclic matching on $\varphi_G^{-1}(v)$ that has $l$ critical simplices of dimension $d+1$, one critical simplex $\tau$ of dimension $1$, and additionally we have the critical simplex that has only the vertex $v$. Finally we match $v$ with $\tau$ and this matching is acyclic since $v$ is the smallest element of $\varphi_G^{-1}(v)$.
\end{proof}
\begin{prop}
\label{rdthm}
Let $n\geq3$ and $G\subset S_1\times S_{n-1}$ be a subgroup, then the topological space $\Delta(\overline{\Pi}_n)/G$ is homotopy equivalent to a wedge of spheres of dimension $n-3$.
\end{prop}
\begin{proof}
We show there exists an acyclic matching on the poset ${\cal F}(\Delta(\overline{\Pi}_n)/G)$, where the set of critical simplices consists of one critical simplex of dimension $0$ and the other critical simplices have dimension $n-3$. Then we can apply Theorem \ref{morsemain}.

For $n=3$ we have the cases $G=S_1\times S_2$ and $G=\{\id_{[3]}\}$, hence $\Delta(\overline{\Pi}_n)/G$ consists of two or three points. We do not match anything and get an acyclic matching with one critical simplex of dimension $0$ and another one or two critical simplices of dimension $0$. Now we assume $n>3$ and proceed by induction.

By the discussion at the end of section \ref{allgemein} it remains to show that there exists an acyclic matching on $\varphi_G^{-1}(v)$ for each $v\in V_G$ such that all critical simplices have dimension $n-3$. Let $v\in V_G$, then we choose a $2\leq k\leq n$ such that $v_k$ is a representative of $v$. By induction hypothesis we have an acyclic matching on $\Delta(\overline{\Pi}_{n-1})/i_k^{-1}(G)$ which has one critical simplex of dimension $0$ and critical simplices of dimension $n-4$. We apply Lemma \ref{indlemma} and get our acyclic matching on $\varphi_G^{-1}(v)$.
\end{proof}
\section{Counting spheres}
\label{giacomo}
We are now interested in the number of spheres we have. This can be computed either by looking into the proof of Proposition \ref{rdthm} or by determining the Betti Number $\beta_{n-3}(\Delta(\overline{\Pi}_n)/G)$, which is the dimension of $H_{n-3}(\Delta(\overline{\Pi}_n)/G;\C)$.

The action of $S_1\times S_{n-1}$ on $\Delta(\overline{\Pi}_n)$ induces an action on $H_{n-3}(\Delta(\overline{\Pi}_n);\C)$. By applying Corollary 3.3 in \cite{emk} and Theorem 1.5 in \cite{white} we know that $H_{n-3}(\Delta(\overline{\Pi}_n);\C)$ and
\[\C\cdot(S_1\times S_{n-1})=\bigoplus_{\sigma\in S_1\times S_{n-1}}\C\cdot\sigma\]
are isomorphic as $(S_1\times S_{n-1})$-modules. The Transfer Theorem tells us
\[
H_{n-3}(\Delta(\overline{\Pi}_n)/G;\C)\cong H_{n-3}(\Delta(\overline{\Pi}_n);\C)^G\cong(\C\cdot(S_1\times S_{n-1}))^G
\]
The dimension of $(\C\cdot(S_1\times S_{n-1}))^G$ is the index of $G$ in $S_1\times S_{n-1}$. Hence we obtain the following result: $\Delta(\overline{\Pi}_n)/G$ is homotopy equivalent to a wedge of $k$ spheres of dimension $n-3$, where $k$ is the index of $G$ in $S_1\times S_{n-1}$.
\section*{Acknowledgments}
The author would like to thank Dmitry N. Feichtner-Kozlov for this interesting problem and the support of this work, Juliane Lehmann for the helpful discussions and Giacomo d'Antonio for the hint that led to an improvement of section \ref{giacomo}.
\bibliographystyle{model1-num-names}
\begin{thebibliography}{00}
\bibitem{transfer}
G.E. Bredon, \textit{Introduction to compact transformation groups}, Pure and applied mathematics \textbf{46}, Chapter III, Academic Press, 1972.
\bibitem{emk}
E.M. Feichtner, \textit{Complexes of trees and nested set complexes}, Pacific journal of mathematics \textbf{227} (2) (2006), pp. 271-286.
\bibitem{forman}
R. Forman, \textit{Morse theory for cell complexes}, Adv. Math. \textbf{134} (1) (1998), pp. 90-145.
\bibitem{braid1}
M. Goresky, R. MacPherson, \textit{Stratified Morse theory}, Springer-Verlag, Berlin Heidelberg New York, 1988.
\bibitem{hatcher}
A. Hatcher, \textit{Algebraic Topology}, Chapter 2, Cambridge University Press, 2008.
\bibitem{clmap}
D.N. Kozlov, \textit{Closure maps on regular trisps}, Topology and its Applications \textbf{156} (15) (2009), pp. 2491-2495.
\bibitem{buch}
D.N. Kozlov, \textit{Combinatorial Algebraic Topology}, Algorithms and Computation in Mathematics \textbf{21}, Springer-Verlag Berlin Heidelberg, 2008.
\bibitem{sing1}
D.N. Kozlov, \textit{Topology of spaces of hyperbolic polynomials and combinatorics of resonances}, Israel Journal of Mathematics \textbf{132} (2002), pp. 189-206.
\bibitem{sing2}
D.N. Kozlov, \textit{Rational homology of spaces of complex monic polynomials with multiple roots}, Mathematika  \textbf{49} (2002), pp. 77-91.
\bibitem{jule}
J. Lehmann, \textit{Equivariant closure operators and trisp closure maps}, Topology and its Applications \textbf{157} (7) (2010), pp. 1195-1201.
\bibitem{braid2}
P. Orlik, H. Terao, \textit{Arrangements of Hyperplanes}, A Series of Comprehensive Studies in Mathematics \textbf{300}, Springer-Verlag, 1992.
\bibitem{white}
A. Robinson, S. Whitehouse, \textit{The tree representation of $\Sigma_{n+1}$}, Journal of Pure and Applied Algebra \textbf{111} (1996), pp. 245-253.
\bibitem{sing3}
V.A. Vassiliev, \textit{Complements of Discriminants of Smooth Maps: Topology and its Applications}, Translations of Mathematical Monographs \textbf{98}, American Mathematical Society, 1994.
\end{thebibliography}
\end{document}

### 2015-09-11 paper2/paper.tex
Sourcefiles for the first 2 papers

\documentclass{elsarticle}
\usepackage{amsmath,amsthm}
\usepackage{bbm}
\usepackage[all,cmtip]{xy}
\usepackage{graphicx}
\usepackage{color}
\newtheorem{df}{Definition}[section]
\newtheorem{prop}[df]{Proposition}
\newtheorem{lem}[df]{Lemma}
\newtheorem{thm}[df]{Theorem}
\newtheorem{rem}[df]{Remark}
\def\N{\mathbbm N}
\def\Z{\mathbbm Z}
\def\R{\mathbbm R}
\def\id{\textrm{id}}
\def\im{\textrm{im}}
\journal{Topology and its Applications}
\begin{document}
\begin{frontmatter}
\title{On a quotient topology of the partition lattice with forbidden block sizes}
\author[ruelle]{Ralf Donau}
\address[ruelle]{Fachbereich Mathematik, Universit\"at Bremen, Bibliothekstra\ss e 1, 28359 Bremen, Germany}
\ead{ruelle@math.uni-bremen.de}
\begin{abstract}
In this paper we consider subtrisps of $\Delta(\Pi_n)/S_1\times S_{n-1}$, that we obtain by forbidding certain block sizes, and determine their homotopy type, as well as bases of their cohomology. Our arguments make use of trisp closure maps, which were introduced by Kozlov. A similar result, where $\Delta(\Pi_n)/S_n$ is considered, has already been solved by Kozlov.
\end{abstract}
\begin{keyword}
Discrete Morse Theory\sep Regular trisp\sep Young subgroup\sep Trisp closure map
\end{keyword}
\end{frontmatter}
\section{Introduction}
Let $n>k\geq 2$ and let $\Pi_n$ denote the poset consisting of all partitions of $[n]:=\{1,\dots,n\}$ ordered by refinement, such that the finer partition is the smaller partition. Let $\Pi_{n,k}$ denote the poset obtained from $\Pi_n$ by removing both the smallest and greatest element, which are $\{\{1\},\dots,\{n\}\}$ and $\{[n]\}$, respectively, and all partitions where some block has more than $1$ and less than $k$ elements. We consider $\Pi_{n,k}$ as a category, which is acyclic, and define $\Delta(\Pi_{n,k})$ to the nerve of the acyclic category $\Pi_{n,k}$, which is a regular trisp, see \cite[Chapter 10]{buch}. Since $\Pi_{n,k}$ is a geometric lattice, $\Delta(\Pi_{n,k})$ is shellable, hence homotopy equivalent to a wedge of spheres, see \cite{bjoern}.

The symmetric group $S_n$ operates on $\Delta(\Pi_{n,k})$ in a natural way. We consider quotients $\Delta(\Pi_{n,k})/G$, where $G\subset S_n$ is a subgroup. The following theorem is the first result concerning the topology of such a quotient, where $G$ is a non-trivial subgroup of $S_n$.
\begin{thm}[Kozlov \cite{kozthm}]
Let $n>k\geq 2$, then $\Delta(\Pi_{n,k})/S_n$ is collapsible.
\end{thm}
In this paper we study the case, where $G$ is the Young subgroup $S_1\times S_{n-1}:=\{\sigma\in S_n\mid\sigma(1)=1\}$. We will use the following result, which gives an answer for the special case $k=2$, in the proof of our main result which is stated in Proposition \ref{mainresult}.
\begin{thm}[Donau \cite{donau}]
\label{rdthm}
Let $n\geq3$, then the topological space $\Delta(\Pi_{n,2})/S_1\times S_{n-1}$ is homotopy equivalent to a sphere of dimension $n-3$.
\end{thm}
A more general version of Theorem \ref{rdthm}, where arbitrary subgroups of $S_1\times S_{n-1}$ are considered, can be found in \cite{donau}.
\section{Closure maps on regular trisps and Discrete Morse Theory}
The definitions of regular trisps, acyclic matchings and foundations of Discrete Morse Theory can be found in \cite{forman}, \cite{clmap} and \cite{buch}.
\begin{thm}%[Main Theorem of Discrete Morse Theory for CW Complexes]
\label{morsemain}
Let $\Delta$ be a regular trisp and let $M$ be an acyclic matching on the poset ${\cal F}(\Delta)\setminus\{\hat{0}\}$. If the critical simplices form a subcomplex $\Delta_c$ of $\Delta$, then $\Delta$ collapses onto $\Delta_c$.
\end{thm}
The proof of Theorem \ref{morsemain} as well as further facts on Discrete Morse Theory can be found in \cite[Chapter 11]{buch}.

For a trisp $\Delta$ we denote the face poset by ${\cal F}(\Delta)$ and by $V(\Delta)$ we denote the set of vertices of $\Delta$. For a simplex $\sigma\in\Delta$, $V(\sigma)$ denotes the set of vertices of $\sigma$.

Let $\Delta$ be a regular trisp whose set of vertices is divided into two disjoint parts $B\cup R$. We call the vertices of $B$ the \emph{blue} vertices and the vertices of $R$ the \emph{red} vertices. Let $\Delta^R$ be the subtrisp of $\Delta$ that contains all simplices $\sigma\in{\cal F}(\Delta^R)$ with $V(\sigma)\subset R$. For $\sigma\in{\cal F}(\Delta)\setminus{\cal F}(\Delta^R)$ let $m(\sigma)$ denote the smallest blue vertex of $\sigma$.
\begin{df}
A map $\varphi:B\longrightarrow R$ is called a \emph{closure map} if for any $\sigma\in{\cal F}(\Delta)\setminus{\cal F}(\Delta^R)$ either $m(\sigma)\in V(\sigma)$ or $m(\sigma)$ can be uniquely inserted into $\sigma$, i.e. there exists a unique simplex $\tau\in{\cal F}(\Delta)$ with $m(\sigma)\in V(\tau)$ and $\sigma$ is a subsimplex of $\tau$ with $\dim\tau=\dim\sigma+1$.
\end{df}
\begin{thm}[Kozlov \cite{clmap}]
\label{clthm}
Let $\Delta$ be a regular trisp whose set of vertices is divided into two disjoint parts $B\cup R$. Assume we have a closure map $\varphi:B\longrightarrow R$, then there exists an acyclic matching on ${\cal F}(\Delta)$ such that a simplex $\sigma\in{\cal F}(\Delta)$ is critical if and only if $\sigma\in{\cal F}(\Delta^R)$. In particular $\Delta$ collapses onto its subtrisp $\Delta^R$.
\end{thm}
\begin{rem}
\label{coll}
If a trisp $\Delta$ collapses onto a subtrisp $\Delta'$, then $\Delta$ is homotopy equivalent to $\Delta'$. If $\Delta'$ consists of one vertex, we call $\Delta$ \emph{collapsible}.
\end{rem}
\section{The main result}
Let $n\geq 3$ and $2\leq k<n$ be a fixed natural numbers throughout this section.

The simplices of $\Delta(\Pi_{n,k})/S_1\times S_{n-1}$ are represented by sequences of partitions, where the partitions refine each other. Such a sequence can be considered as a leveled forest, where each level corresponds to a partition. Each node on a particular level corresponds to a block of the corresponding partition in the sequence. Now we modify such a forest as follows. We replace each node, which is a subset of $[n]$, by its cardinality and put a mark onto the number if this subset contains $1$. It is easy to see that the simplices of $\Delta(\Pi_{n,k})/S_1\times S_{n-1}$ can be indexed with such forests. The vertices can be indexed with number partitions, which we may write as $v_0\oplus v_1+\dots+v_r$, of $n$ that distinguish the first number, i.e. $\oplus$ is non-commutative. The number on the left side of $\oplus$, that is $v_0$, corresponds to the block that contains $1$. We call this number the \emph{number on the left side}. All other numbers are on the right side. The boundary operators are obtained by deleting entire levels from forests and reconnecting vertices transitively through the deleted levels.
\begin{figure}[ht]
\centering
\input{forest.pdf_t} 
\caption{A simplex of $\Delta(\Pi_{6,3})/S_1\times S_5$. One number is distinguished on each level.}
\end{figure}

Let $\Delta_{n,k}$ denote the subtrisp of $\Delta(\Pi_{n,k})/S_1\times S_{n-1}$ where all vertices are indexed with number partitions of the type
\[v_0\oplus k^c+1^{n-v_0-ck}:=v_0\oplus\underbrace{k+\dots+k}_\text{$c$ times}+\underbrace{1+\dots+1}_\text{$n-v_0-ck$ times}\]
with $v_0\equiv 0,1\mod k$.
\begin{lem}
\label{erste_closure}
The trisp $\Delta(\Pi_{n,k})/S_1\times S_{n-1}$ is homotopy equivalent to $\Delta_{n,k}$.
\end{lem}
Some parts of the argument are similar to the proof of Corollary 2.3 in \cite{clmap}.
\begin{proof}
We call the vertices of $\Delta_{n,k}$ the red vertices of $\Delta(\Pi_{n,k})/S_1\times S_{n-1}$. Let $R$ denote the set of the red vertices and $B$ the set of all other vertices of $\Delta$ that are blue. We define a closure map $\varphi:B\longrightarrow R$ as follows. Let $b=b_0\oplus b_1+\dots+b_r\in B$ be a blue vertex. We refine the blocks of $b$ as follows. $b_0$ remains unchanged for $b_0\equiv 0,1\mod k$. If we have $b_0\equiv a\mod k$ with $a=2,\dots,k-1$, we replace the block $b_0$ by the blocks $(b_0-a+1)\oplus1^{a-1}$. For $i\not=0$, each $b_i$ is replaced as follows. We write $b_i$ as a sum of $k$'s and $1$'s such that we maximize the number of $k$'s.

We have to show that $\varphi$ is a closure map. Let $\sigma$ be a simplex with smallest blue vertex $m(\sigma)=m_0\oplus m_1+\dots+m_r$ and assume $\varphi(m(\sigma))$ is not a vertex of $\sigma$. We have to construct a forest $\tau$ such that $\varphi(m(\sigma))$ is a vertex of $\tau$ and $\sigma$ can be obtained from $\tau$ by deleting a level, in our case this is the vertex $\varphi(m(\sigma))$. If $m(\sigma)$ is the smallest vertex of $\sigma$, then clearly $\varphi(m(\sigma))$ can be appended uniquely below $m(\sigma)$. Assume we have red vertices $r^1,\dots,r^t\in V(\sigma)$ that are smaller than $m(\sigma)$, where $r^t$ is the greatest vertex with this property. We start with the subsimplex of $\sigma$ that we obtain by deleting the vertices $r^1,\dots,r^t$. Then, as we have seen above, $\varphi(m(\sigma))$ can be appended below $m(\sigma)$. Now we append $r^t$ below $\varphi(m(\sigma))$ as follows. First we consider the number $m_0$ on the left side of $m(\sigma)$. If we have $b_0\equiv a\mod k$ with $a=2,\dots,k-1$, then this number has exactly $a-1$ children which are $1$'s on the right side of $\varphi(m(\sigma))$. If we look at $\sigma$ we see that this number has at least $a-1$ children which are $1$'s on the right side of $r^t$. We let $a-1$ of these $1$'s in $r$ become children of the corresponding $1$'s in $\varphi(m(\sigma))$ and the other $1$'s become children of the number on the left side of $\varphi(m(\sigma))$. For $b_0\equiv 0,1\mod k$, the number $m_0$ has exactly one child. We let the children of $m_0$ become the children of the number on the left side of $\varphi(m(\sigma))$. Now we let $m_i$ be a number on the right side of $m(\sigma)$. In $r^t$, this number is refined into $k$'s and $1$'s but we do not necessarily have maximized the number of $k$'s. These numbers become children of the corresponding number in $\varphi(m(\sigma))$ where the number of $k$'s is maximized. At the end we append the remaining children $r^1,\dots,r^{t-1}$ in the same way as in $\sigma$. The uniqueness can be easily verified. Now we can apply Theorem \ref{clthm} and Remark \ref{coll}.
\end{proof}
\begin{lem}
\label{rdanwendung}
The trisp $\Delta_{n,2}$ is homotopy equivalent to a sphere of dimension $n-3$.
\end{lem}
\begin{proof}
Apply Lemma \ref{erste_closure} and Theorem \ref{rdthm}.
\end{proof}
\begin{figure}[ht]
\centering
\input{S1.pdf_t}
\caption{$\Delta_{4,2}$ as a circle, where $v_1=2\oplus2$, $v_2=1\oplus 2+1$, $v_3=2\oplus 1 +1$, $v_4=3\oplus 1$.}
\end{figure}

Now assume $n\equiv 0,1\mod k$. Each vertex of $\Delta_{n,k}$ can be written as
\[(ak+b)\oplus k^c+1^{n-ak-b-ck}\]
with $b\in\{0,1\}$, $a,b,c\in\Z$, $a,b,c\geq 0$, $a+b>0$, $a+c>0$, $ak+b+ck\leq n$ and $ak+b<n$ such that $(a,b,c)$ is unique. The two last inequalities can be rewritten as $0<a+c\leq\frac nk-\frac bk$ and $a<\frac nk-\frac bk$. Now we would like to eliminate $\frac bk$ from these inequalities. For $n\equiv 0\mod k$ we have
\begin{eqnarray*}
a+c\leq\frac nk-\frac bk&\Leftrightarrow&a+c\leq\frac nk-b\\
a<\frac nk-\frac bk&\Leftrightarrow&a<\frac nk
\end{eqnarray*}
and for $n\equiv 1\mod k$ we have
\begin{eqnarray*}
a+c\leq\frac nk-\frac bk&\Leftrightarrow&a+c\leq\frac {n-1}k\\
a<\frac nk-\frac bk&\Leftrightarrow&a<\frac {n-1}k+1-b
\end{eqnarray*}
Hence in the case $n\equiv 0\mod k$ the set of tuples $(a,b,c)$ only depends on $\frac nk$ and in the case $n\equiv 1\mod k$ this set only depends on $\frac {n-1}k$. In particular the number of vertices of $\Delta_{n,k}$ only depends on these numbers.
\begin{lem}
\label{trispiso}
Let $n_i\geq3$ and $2\leq k_i<n_i$ for $i=1,2$. Assume $n_i\equiv 0\mod k_i$ or $n_i\equiv 1\mod k_i$, and assume $\frac{n_1}{k_1}=\frac{n_2}{k_2}$ or $\frac{n_1-1}{k_1}=\frac{n_2-1}{k_2}$, respectively. Then the trisps $\Delta_{n_1,k_1}$ and $\Delta_{n_2,k_2}$ are isomorphic, in particular homotopy equivalent.
\end{lem}
\begin{proof}
We obtain the canonical bijection
\begin{eqnarray*}
\varphi:V(\Delta_{n_1,k_1})&\longrightarrow&V(\Delta_{n_2,k_2})\\
(ak_1+b)\oplus k_1^c+1^{n_1-ak_1-b-ck_1}&\longmapsto&(ak_2+b)\oplus k_2^c+1^{n_2-ak_2-b-ck_2}
\end{eqnarray*}
We also obtain a bijection between ${\cal F}(\Delta_{n_1,k_1})$ and ${\cal F}(\Delta_{n_2,k_2})$ as follows. Let $\sigma\in{\cal F}(\Delta_{n_1,k_1})$ be a simplex, which is a forest. First, we remove all $1$'s from the tree which are blocks on the right side. Second, on the right side we replace each $k_1$ by $k_2$, on the left side we replace $ak_1+b$ by $ak_2+b$. The last step is: We add $1$'s to each level such that the sum over all numbers is $n_2$ and such that they become children of those blocks which do not have enough children. The last step is well-defined. It is easy to see that $\varphi$ is a trisp isomorphism.
\begin{figure}[ht]
\centering
\input{mapping.pdf_t} 
\caption{Example for the trisp isomorphism}
\end{figure}
\end{proof}
\begin{prop}
\label{mainresult}
Let $n\geq3$ and $2\leq k<n$, then the topological space $\Delta(\Pi_{n,k})/S_1\times S_{n-1}$ is
\begin{enumerate}
\item collapsible for $n\not\equiv 0,1\mod k$.
\item homotopy equivalent to a sphere of dimension $\frac{2n}k-3$ for $n\equiv 0\mod k$.
\item homotopy equivalent to a sphere of dimension $\frac{2(n-1)}k-2$ for $n\equiv 1\mod k$.
\end{enumerate}
\end{prop}
\begin{proof}
Since $\Delta(\Pi_{n,k})/S_1\times S_{n-1}$ is homotopy equivalent to $\Delta_{n,k}$ by Lemma \ref{erste_closure}, we may consider $\Delta_{n,k}$.

Assume we have  $n\equiv 2,\dots,k-1\mod k$. Let $v=v_0\oplus k^c+1^{n-v_0-ck}$ be a vertex of $\Delta_{n,k}$, then we have $v_0\equiv 0,1\mod k$. This implies $n-v_0\equiv 1,\dots,k-1\mod k$, in particular $n-v_0-ck\not\equiv 0\mod k$. Hence each vertex has a $1$ on the right side and $(n-1)\oplus 1$ can be uniquely inserted into each simplex of $\Delta_{n,k}$. We define a closure map where $(n-1)\oplus 1$ is the only red vertex and apply Theorem \ref{clthm}.

In the case $n\equiv 0\mod k$ we set $n_2:=\frac{2n}k$, then $\Delta_{n,k}$ is isomorphic to $\Delta_{n_2,2}$ by Lemma \ref{trispiso}, since $\frac nk=\frac{n_2}2$. In the case $n\equiv 1\mod k$ we set $n_2:=\frac{2(n-1)}k+1$, then $\Delta_{n,k}$ is isomorphic to $\Delta_{n_2,2}$ by Lemma \ref{trispiso}, since $\frac {n-1}k=\frac{n_2-1}2$. Lemma \ref{rdanwendung} completes the proof.
\end{proof}
\section{A basis of the cohomology}
Let $n\geq3$, $2\leq k<n$ and $n\equiv 0,1\mod k$. Then $\Delta:=\Delta(\Pi_{n,k})/S_1\times S_{n-1}$ is homotopy equivalent to a sphere of some dimension $l$ by Proposition \ref{mainresult}. We clearly have 
\[
H^i(\Delta)\cong
\begin{cases}
\Z&\text{for $i=0$}\\
\Z&\text{for $i=l$}\\
0&\text{else}
\end{cases}
\]
$H^i(\Delta)$ denotes the $i$th cohomology of $\Delta$ with coefficients in $\Z$. Now we wish to give an element that generates the top cohomology group $H^l(\Delta)$. Let $(C_*,\partial_*)$ denote the chain complex of the trisp $\Delta$ with coefficients in $\Z$, see \cite[Chapter 3]{buch}. Then for each $i>0$, $\partial_i:C_i\longrightarrow C_{i-1}$ is defined as follows. Let $\tau$ be a simplex of dimension $i$, then we have $\partial_i(\tau)=\sum_{i=1}^{i+1}(-1)^{i+1}B_i(\tau)$. Here $B_i(\tau)$ denotes that subsimplex of $\tau$, that is obtained by removing the vertex on the $i$th level. A detailed description of $B_i$ can be found in \cite{buch}.

We consider the cochain complex of $\Delta$ at the $l$th position, see \cite[Section 3.1]{hatcher}:
\[
\xymatrix{
\dots&\hom(C_{l+1},\Z)\ar[l]&\hom(C_l,\Z)\ar[l]_{\partial^*}&\hom(C_{l-1},\Z)\ar[l]_{\partial^{**}}&\dots\ar[l]
}
\]
We give is a short description of $\partial^*$. Let $\tau$ be simplex of dimension $l+1$.
\[
(\partial^*f)(\tau)=\sum_{i=1}^{l+2}(-1)^{i+1}f(B_i(\tau))
\]
$H^l(\Delta)$ is defined as $H^l(\Delta):=\ker\partial^*/\im\partial^{**}$.

We define the following subsets of $V(\Delta)$ which are pairwise disjoint.
\begin{eqnarray*}
S_{n,k}&:=&\left\{ak\oplus 1^{n-ak}\mid 1\leq a<\frac nk\right\}\\
T_{n,k}^a&:=&\{(ak+b)\oplus 1^{n-ak-b}\mid b=1,\dots,k-1\}\\
&\text{for}&1\leq a<\frac {n-1}k
\end{eqnarray*}

We define a homomorphism $\rho_{n,k}\in\hom(C_l,\Z)$ as follows. $C_l$ is generated by the simplices of dimension $l$. Let $\tau\in C_l$ be a simplex of dimension $l$. We set $\rho_{n,k}(\tau)=1$ if $S_{n,k}\subset V(\tau)$ and $T_{n,k}^a\cap V(\tau)\not=\emptyset$ for all $1\leq a<\frac {n-1}k$. We set $\rho_{n,k}(\tau)=0$ otherwise.
\begin{lem}
\label{cocycle}
The homomorphism $\rho_{n,k}$ is a cocycle, i.e. $\partial^*\rho_{n,k}=0$.
\end{lem}
\begin{proof}
Let $\tau\in C_{l+1}$ be a simplex of dimension $l+1$. Assume $S_{n,k}$ is not contained in $V(\tau)$ or  $T_{n,k}^a\cap V(\tau)$ is empty for some $1\leq a<\frac {n-1}k$, then this also holds for each subsimplex of $\tau$, hence $(\partial^*\rho_{n,k})(\tau)=0$.

Now assume $S_{n,k}\subset V(\tau)$ and $T_{n,k}^a\cap V(\tau)\not=\emptyset$ for all $1\leq a<\frac {n-1}k$. Notice that $\tau$ has $l+2$ vertices. We have $|T_{n,k}^a\cap V(\tau)|\leq 2$ for all $1\leq a<\frac {n-1}k$ and $|T_{n,k}^a\cap V(\tau)|=2$ for at most one $a$ by the following calculation. First we consider the case $n\equiv 0\mod k$. Then we have $l=\frac{2n}k-3$ and $|S_{n,k}|=\frac nk-1$.
\[
\frac{2n}k-2=\frac nk-1+\frac nk-1\leq |S_{n,k}|+\sum_{a=1}^{\frac nk-1}|T_{n,k}^a\cap V(\tau)|\leq l+2=\frac{2n}k-1
\]
In the case $n\equiv 1\mod k$ we have $l=\frac{2(n-1)}k-2$ and $|S_{n,k}|=\frac {n-1}k$.
\[
\frac{2(n-1)}k-1=\frac {n-1}k+\frac {n-1}k-1\leq |S_{n,k}|+\sum_{a=1}^{\frac {n-1}k-1}|T_{n,k}^a\cap V(\tau)|\leq l+2=\frac{2(n-1)}k
\]
Now it is easy to see that $(\partial^*\rho_{n,k})(\tau)=0$.
\end{proof}
\begin{lem}
\label{nicematching}
Let $n\geq3$, then there exists an acyclic matching on the poset ${\cal F}(\Delta(\Pi_{n,2})/S_1\times S_{n-1})$, where the set of critical simplices consists of one critical simplex of dimension $0$ and the unique simplex $\sigma_n$ of dimension $n-3$ which has the vertices $v_0\oplus 1^{n-v_0}$ with $v_0=2,\dots,n-1$.
\end{lem}
This proof may be considered as a simplified version of the proof of Proposition 4.2 in \cite{donau}. Lemma \ref{nicematching} implies Theorem \ref{rdthm} by the Main Theorem of Discrete Morse Theory\footnote{See \cite[Chapter 11]{buch}}.
\begin{proof}
For $n=3$ we have the critical simplices $2\oplus1$ and $1\oplus2$ if we do not match anything. Now we assume $n>3$ and proceed by induction.

We define an order-preserving map:
\begin{eqnarray*}
\varphi:{\cal F}(\Delta(\Pi_{n,2})/S_1\times S_{n-1})&\longrightarrow&[0,1]\\
\tau&\longmapsto&
\begin{cases}
1&\text{if $2\oplus 1^{n-2}\in V(\tau)$}\\
0&\text{else}
\end{cases}
\end{eqnarray*}
By Lemma 3.2 in \cite{donau} we already have an acyclic matching on $\varphi^{-1}(0)$ with one critical simplex which has dimension $0$. The next step is to find an acyclic matching on $\varphi^{-1}(1)$ and apply the Patchwork Theorem, see \cite[Chapter 11]{buch}.

We define a map $\psi:{\cal F}(\Delta(\Pi_{n-1,2})/S_1\times S_{n-2})\longrightarrow\varphi^{-1}(1)\setminus\{2\oplus 1^{n-2}\}$, which increases the number on the left side by $1$ on each level and appends the vertex $2\oplus 1^{n-2}$ at the bottom of the forest. The map $\psi$ is an isomorphism between posets, see Lemma 4.1 in \cite{donau}, and $\psi$ induces an acyclic matching on $\varphi^{-1}(1)$ where the set of critical simplices consists of one simplex $\alpha$ of dimension $1$, the simplex that has only the vertex $2\oplus 1^{n-2}$ and the unique simplex $\sigma_n$ with the vertices $v_0\oplus 1^{n-v_0}$ with $v_0=2,\dots,n$ by induction hypothesis. Finally we match $2\oplus 1^{n-2}$ with $\alpha$.
\end{proof}
It is clear that $H^{n-3}(\Delta(\Pi_{n,2})/S_1\times S_{n-1})$ is generated by an element that is represented by a function $C_{n-3}(\Delta(\Pi_{n,2})/S_1\times S_{n-1})\longrightarrow\Z$ that maps the simplex $\sigma_n$ to $1$ and any other simplex of dimension $n-3$ to $0$. Furthermore, since $\sigma_n$ is also a simplex of $\Delta_{n,2}$, $H^{n-3}(\Delta_{n,2})$ is generated by the class of a function $\delta_{\sigma_n}:C_{n-3}(\Delta_{n,2})\longrightarrow\Z$ that maps $\sigma_n$ to $1$ and any other simplex to $0$.

Via the map in the proof of Lemma \ref{trispiso} we obtain the unique simplex $\sigma_{n,k}$ of dimension $l$ with the vertices $v_0\oplus 1^{n-v_0}$ with $v_0=2,\dots,n-1$ and $v_0\equiv 0,1\mod k$, such that $\{[\delta_{\sigma_{n,k}}]\}$ is a basis of the free group $H^l(\Delta_{n,k})$. $[\delta_{\sigma_{n,k}}]$ denotes the class of $\delta_{\sigma_{n,k}}$.
\begin{prop}
The set $\{[\rho_{n,k}]\}$ is a basis of the free group $H^l(\Delta(\Pi_{n,k})/S_1\times S_{n-1})$.
\end{prop}
\begin{proof}
By Lemma \ref{cocycle}, $\rho_{n,k}$ is a cocycle. The homomorphism $\rho_{n,k}$ is not a coboundary by the following argument: Assume there exists $g\in C^{l-1}$ with $\rho_{n,k}=\partial^*g$. If we restrict $\rho_{n,k}$ to $\Delta_{n,k}$, this implies $\sigma_{n_k}=\partial^*\widetilde g$, where $\widetilde g$ is the restriction of $g$ to $\Delta_{n,k}$. It remains to show that $\rho_{n,k}$ is a generator. Assume there exists $f\in\hom(C_l,\Z)$ such that $[\rho_{n,k}]=k[f]$ with $k>1$. Again we restrict to $\Delta_{n,k}$ and obtain $[\sigma_{n,k}]=k[\widetilde f]$. By the discussion above this is a contradiction.
\end{proof}
%\section*{Acknowledgements}
%The author would like to thank ...
\bibliographystyle{model1-num-names}
\begin{thebibliography}{00}
\bibitem{bjoern}
A. Bj\"orner, \textit{Matroid Applications}, Cambridge University Press, 1992.
\bibitem{donau}
R. Donau, \textit{Quotients of the order complex $\Delta(\overline{\Pi}_n)$ by subgroups of the Young subgroup $S_1\times S_{n-1}$}, Topology and its Applications \textbf{157} (16) (2010), pp. 2476-2479.
\bibitem{forman}
R. Forman, \textit{Morse theory for cell complexes}, Adv. Math. \textbf{134} (1) (1998), pp. 90-145.
\bibitem{hatcher}
A. Hatcher, \textit{Algebraic Topology}, Cambridge University Press, 2008.
\bibitem{clmap}
D.N. Kozlov, \textit{Closure maps on regular trisps}, Topology and its Applications \textbf{156} (15) (2009), pp. 2491-2495.
\bibitem{kozthm}
D.N. Kozlov, \textit{Collapsibility of $\Delta(\Pi_n)/S_n$ and some related CW complexes}, Proc. Amer. Math. Soc. \textbf{128} (8) (2000), pp. 2253-2259.
\bibitem{buch}
D.N. Kozlov, \textit{Combinatorial Algebraic Topology}, Algorithms and Computation in Mathematics \textbf{21}, Springer-Verlag Berlin Heidelberg, 2008.
\end{thebibliography}
\end{document}

### 2018-04-25 paper4/eqalg.tex
Algebraic

\documentclass{elsarticle}
\usepackage{amsmath,amsthm}
\usepackage{bbm}
\newtheorem{df}{Definition}[section]
\newtheorem{prop}[df]{Proposition}
\newtheorem{lem}[df]{Lemma}
%\newtheorem{thm}[df]{Theorem}
\newtheorem{cor}[df]{Corollary}
%\newtheorem{ex}[df]{Example}
\newtheorem{rem}[df]{Remark}
\def\Z{\mathbbm Z}
\def\id{\operatorname{id}}
\def\im{\operatorname{im}}
\def\span{\operatorname{span}}

\journal{arXiv.org}
\begin{document}
\begin{frontmatter}
\title{Equivariant Algebraic Morse Theory}
\author[ruelle]{Ralf Donau}
\address[ruelle]{Fachbereich Mathematik, Universit\"at Bremen, Bibliothekstra\ss e 1, 28359 Bremen, Germany}
\ead{ruelle@math.uni-bremen.de}
\begin{abstract}
In this paper we develop Algebraic Morse Theory for the case where a group acts on a free chain complex. Algebraic Morse Theory is an adaption of Discrete Morse Theory to free chain complexes.
\end{abstract}
\begin{keyword}
Discrete Morse Theory\sep Acyclic matching\sep Algebraic Morse Theory
\end{keyword}
\end{frontmatter}
\section{Introduction}
There exists an equivariant version of the Main Theorem of Discrete Morse Theory, see \cite{freij}. In this paper I present an equivariant version of Theorem 11.24 in \cite[Chapter 11.3]{buch}. We use the same notion of an equivariant acyclic matching as in Equivariant Discrete Morse Theory. An example for working with equivariant acyclic matchings can be found in \cite{donau3}.
\section{Equivariant acyclic matchings}
The definition of an acyclic matching on a poset can be found in \cite{clmap,buch}.
\begin{df}
Let $P$ be a poset and let $G$ be a group acting on $P$. Let $M$ be an acyclic matching on $P$. We call $M$ an \emph{$G$-equivariant acyclic matching} if $(a,b)\in M$ implies $(ga,gb)\in M$ for all $g\in G$ and $a,b\in P$.
\end{df}
There exists a characterization of acyclic matchings by means of order-preserving maps with small fibers, see Definition 11.3 and Theorem 11.4 in \cite[Chapter 11]{buch}. In a similar way we can also characterize $G$-equivariant acyclic matchings by means of order-preserving $G$-maps with small fibers.

For an order-preserving map with small fibers $\varphi$ let $M(\varphi)$ denote its associated acyclic matching which consists of all fibers of cardinality $2$, see \cite[Chapter 11]{buch}.
\begin{prop}
\label{smallfibers}
Let $G$ be a group acting on a finite poset $P$. For any order-preserving $G$-map $\varphi:P\longrightarrow Q$ with small fibers, the acyclic matching $M(\varphi)$ is $G$-equivariant. On the other hand, any $G$-equivariant acyclic matching $M$ on $P$ can be represented as $M=M(\varphi)$, where $\varphi:P\longrightarrow Q$ is an order-preserving $G$-map with small fibers.
\end{prop}
The proof of Proposition \ref{smallfibers} can be found in \cite{donau3}.
\section{The main result}
We consider chain complexes of modules over some fixed commutative ring $R$ with unit. Furthermore we consider group actions on such chain complexes. Let $G$ be a group and let $C_*=(\dots\overset{\partial_{n+2}}\longrightarrow C_{n+1}\overset{\partial_{n+1}}\longrightarrow C_n\overset{\partial_n}\longrightarrow\dots)$ be a finitely generated free chain complex with an action of the group $G$ and let $\Omega=(\Omega_n)_n$ be a $G$-basis of $C_*$, i.e. each $\Omega_n$ is closed under the action of $G$. For $b\in\Omega_n$ let $k_b:C_n\longrightarrow R$, $\sum_{b'\in\Omega_n}\lambda_{b'}b'\longmapsto\lambda_b$ denote the linear function which maps any $x\in C_n$ to the coefficient of $b$ inside the linear representation of $x$.

Let $P(C_*,\Omega):=\bigcup_n\Omega_n$. We define an order relation on $P(C_*,\Omega)$ as follows. For $a\in\Omega_n$ and $b\in\Omega_{n+1}$ we denote the \emph{weight of the covering relation} by $w(b\succ a):=k_a(\partial_n b)$, we set $a\leq b$ if $w(b\succ a)\not=0$. Furthermore for $a\in\Omega_n$ and $b\in\Omega_m$ with $n\leq m$, we set $a\leq b$ if there exists a sequence $(c_i)_{n\leq i\leq m}$ with $c_i\leq c_{i+1}$ for $n\leq i<m$ such that $a=c_n$ and $b=c_m$. This defines a partial order relation on $P(C_*,\Omega)$, which can be easily verified. Notice that $P(C_*,\Omega)$ is a $G$-poset since $\partial_n$ is a $G$-map.

Let $M$ be an $G$-equivariant acyclic matching on $P(C_*,\Omega)$ such that $w(b\succ a)$ is invertible for any $(a,b)\in M$. Let $\varphi$ denote the order-preserving $G$-map with small fibers with $M(\varphi)=M$, which exists by Proposition \ref{smallfibers}.
\begin{df}
For $b\in C_n$ we define the $G$-subcomplex ${\cal A}(Gb)$ as follows.
\[\dots\longrightarrow 0\longrightarrow\span(Gb)\overset{\partial^{Gb}_n}\longrightarrow\span(G\partial_n b)\longrightarrow 0\longrightarrow\dots\]
$\partial^{Gb}_n$ denotes the restriction of $\partial_n$ to $\span(Gb)$, i.e. $\partial^{Gb}_n(x):=\partial_n(x)$.
\end{df}
Notice that $\partial^{Gb}_n$ is surjective by construction.
\begin{rem}
\label{poset_orbit}
Let $G$ be a finite group acting on a finite poset $Q$. Let $q\in Q$ and $g\in G$. Then $gq\leq q$ implies $gq=q$. In other words, the elements inside an orbit are not comparable to each other.
\end{rem}
Let $(a,b)\in M$ be a matching pair. Then for any $g\in G$, $a\prec gb$ implies $\varphi(b)=\varphi(a)\leq\varphi(gb)=g\varphi(b)$ which implies $\varphi(b)=\varphi(gb)$ by Remark \ref{poset_orbit}. Hence $b=gb$ since $\varphi$ has small fibers. In particular $k_a(\partial gb)=0$ for $gb\not=b$.
\begin{lem}
\label{iso}
Let $(a,b)\in M$ be a matching pair, assume $b\in C_n$. Then $\partial^{Gb}_n$ is an isomorphism and ${\cal A}(Gb)$ is $G$-homotopy equivalent to the zero complex.
\end{lem}
\begin{proof}
We have to show that $\partial^{Gb}_n$ is injective. Assume $\partial_n(\sum_{b'\in Gb}\lambda_{b'}b')=0$. Then $0=k_a(\sum_{b'\in Gb}\lambda_{b'}\partial_nb')=\lambda_{b}k_a(\partial_nb)$, which implies $\lambda_b=0$, since $k_a(\partial_nb)$ is invertible. This implies $\lambda_{b'}=0$ for all $b'\in Gb$, since $\partial_n$ is a $G$-map.

The composition ${\cal A}(Gb)\longrightarrow0\longrightarrow{\cal A}(Gb)$ is homotop to $\id:{\cal A}(Gb)\longrightarrow{\cal A}(Gb)$ via the $G$-chain homotopy
\begin{eqnarray*}
P:{\cal A}(Gb)&\longrightarrow&{\cal A}(Gb)\\
x&\longmapsto&(\partial^{Gb}_n)^{-1}(x)\text{ for $x\in\span(G\partial_n b)$}\\
x&\longmapsto&0\text{ else}
\end{eqnarray*}
On the other hand $0\longrightarrow{\cal A}(Gb)\longrightarrow 0$ equals $\id:0\longrightarrow 0$. Hence ${\cal A}(Gb)$ is $G$-homotopy equivalent to the zero complex.
\end{proof}
\begin{prop}
\label{eqthm}
Let $M$ be an $G$-equivariant acyclic matching on $P(C_*,\Omega)$ such that $w(b\succ a)$ is invertible. Then there exists a $G$-chain complex $C_*^M$ such that $C_*\cong_G C_*^M\oplus T_*$ where $T_*=\bigoplus_{G(a,b)\in M/G}{\cal A}(Gb)$.
\end{prop}
This in an $G$-equivariant version of Theorem 11.24 in \cite[Chapter 11.3]{buch}. Notice that $C_*^M$ is unique up to $G$-isomorphism.
\begin{proof}
By induction on the number $m=|M/G|$ of orbits in $M$. For $m=0$, we set $C_*^M:=C_*$. Now assume $m>0$. By Proposition \ref{smallfibers}, there exists a finite poset $Q$ and an order-preserving $G$-map with small fibers $\varphi:P(C_*,\Omega)\longrightarrow Q$ such that $M(\varphi)=M$. We consider the following subposet of $Q$:
\[
Q':=\{q\in Q\mid\varphi^{-1}(q)=\{a,b\}\text{ with }a\not=b\}
\]
Let $q\in Q'$ be a minimal element. Assume $a\in\Omega_{n-1}$, $b\in\Omega_n$ and $\varphi^{-1}(q)=\{a,b\}$. We construct a new basis $\widetilde\Omega$ by replacing $\Omega_{n-1}$ and $\Omega_n$ as follows. We construct two $G$-automorphisms $f_{n-1}:C_{n-1}\longrightarrow C_{n-1}$ and $f_n:C_n\longrightarrow C_n$.
\begin{eqnarray*}
f_{n-1}:C_{n-1}\supset\Omega_{n-1}&\longrightarrow&C_{n-1}\\
ga&\longmapsto&g\partial_n b\text{ for all $g\in G$}\\
x&\longmapsto&x\text{ for $x\not\in Ga$}
\end{eqnarray*}
We set $\Omega^{Gb}_{n-1}:=\Omega_{n-1}\setminus Ga$, then $\widetilde\Omega_{n-1}:=f_{n-1}(\Omega_{n-1})=\Omega^{Gb}_{n-1}\cup Gf_{n-1}(a)$ generates $C_{n-1}$ since the linear representation of $f_{n-1}(a)$ is
\[
f_{n-1}(a)=\partial_n b=w(b\succ a)\cdot a+\sum_{x\in\Omega^{Gb}_{n-1}}w(b\succ x)\cdot x
\]
which implies
\[
a=\frac 1{w(b\succ a)}\cdot f_{n-1}(a)-\sum_{x\in\Omega^{Gb}_{n-1}}\frac{w(b\succ x)}{w(b\succ a)}\cdot x
\]
hence $a\in\span(f_{n-1}(\Omega_{n-1}))$. This implies $ga\in\span(f_{n-1}(\Omega_{n-1}))$ for all $g\in G$ since $f_{n-1}$ is a $G$-map by construction.

Assume
\[
\sum_{x\in\Omega^{Gb}_{n-1}}\lambda_xx+\sum_{a'\in Ga}\lambda_{a'}f_{n-1}(a')=0
\]
Then in particular
\begin{eqnarray*}
&k_a\left(\sum_{x\in\Omega^{Gb}_{n-1}}\lambda_xx+\sum_{a'\in Ga}\lambda_{a'}f_{n-1}(a')\right)&=0\\
\Longrightarrow&\lambda_a\cdot k_a(\partial_n b)&=0\\
\Longrightarrow&\lambda_a\cdot w(b\succ a)&=0\\
\Longrightarrow&\lambda_a&=0
\end{eqnarray*}
For any $g\in G$ this implies $\lambda_{ga}=0$, because $f_{n-1}$ is a $G$-map. Since $\Omega^{Gb}_{n-1}$ is linear independent, we have $\lambda_x=0$ for all $x\in\Omega^{Gb}_{n-1}$.
\begin{eqnarray*}
f_n:C_n\supset\Omega_n&\longrightarrow&C_n\\
gb&\longmapsto&gb\text{ for all $g\in G$}\\
x&\longmapsto&x-\sum_{b'\in Gb}w(x\succ\partial_n b')\cdot b'\text{ for $x\not\in Gb$}
\end{eqnarray*}
We set $\Omega_n^{Gb}:=f_n(\Omega_n\setminus Gb)$, then $\widetilde\Omega_n:=f_n(\Omega_n)=\Omega_n^{Gb}\cup Gb$ generates $C_n$, since for all $x\in\Omega_n\setminus Gb$ we have
\[
x=f_n(x)+\sum_{b'\in Gb}w(x\succ\partial b')\cdot b'\in\span(\widetilde\Omega_n)
\]
where $f_n(x)\in\span(\Omega_n^{Gb})$ and $\sum_{b'\in Gb}w(x\succ\partial b')\cdot b'\in\span(Gb)$. Clearly we have $k_x(f_n(x))=1$ and $k_x(f_n(x'))=0$ for all $x'\in\Omega_n\setminus\{x\}$. Assume
\[
\sum_{x'\in\Omega_n\setminus Gb}\lambda_{x'} f_n(x')+\sum_{b'\in Gb}\lambda_{b'}b'=0
\]
Then for all $x\in\Omega_n\setminus Gb$:
\begin{eqnarray*}
&k_x\left(\sum_{x'\in\Omega_n\setminus Gb}\lambda_{x'} f_n(x')+\sum_{b'\in Gb}\lambda_{b'}b'\right)&=0\\
\Longrightarrow&\lambda_xk_x(f_n(x))&=0\\
\Longrightarrow&\lambda_x&=0
\end{eqnarray*}
Since $Gb$ is linear independent, this also implies $\lambda_{gb}=0$ for all $g\in G$.

Hence $\widetilde\Omega_{n-1}$ and $\widetilde\Omega_n$ are bases of $C_{n-1}$ and $C_n$ respectively. Let $\widetilde\Omega$ denote the basis of $C_*$ which is obtained from $\Omega$ by replacing $\Omega_{n-1}$ and $\Omega_n$ with these newly constructed bases. Then $\widetilde\Omega$ has the following properties:
\begin{itemize}
 \item[(1)] $\partial_n(\span(\Omega_n^{Gb}))\subset\span(\Omega_{n-1}^{Gb})$
 \item[(2)] $\im\partial_{n+1}\subset\span(\Omega_n^{Gb})$
\end{itemize}
For $b\in\widetilde\Omega$ let $\widetilde k_b$ denote the linear function which maps an element to the coefficient of $b$ inside its linear representation.

For all $x\in\Omega_n\setminus Gb$ we have
\begin{eqnarray*}
\widetilde k_{f_{n-1}(a)}(\partial_n f_n(x))&=&\widetilde k_{f_{n-1}(a)}(\partial_n x)-\sum_{b'\in Gb}w(x\succ\partial_n b')\widetilde k_{f_{n-1}(a)}(\partial_n b')\\
&=&\widetilde k_{f_{n-1}(a)}(\partial_n x)-w(x\succ\partial_n b)\\
&=&\widetilde k_{f_{n-1}(a)}(\partial_n x)-w(x\succ f_{n-1}(a))\\
&=&0
\end{eqnarray*}
Since $f_n$ is a $G$-map, we also have $\widetilde k_{f_{n-1}(ga)}(\partial_n f_n(x))=0$ for all $g\in G$ and all $x\in\Omega_n\setminus Gb$. Hence (1) is proven.

We have $\widetilde k_{f_{n-1}(a)}(\partial_n f_n(b'))=\widetilde k_{f_n(b)}(f_n(b'))$ for all $b'\in Gb$ by construction of $f_{n-1}$, on the other hand we have $0=\widetilde k_{f_{n-1}(a)}(\partial_n x)=\widetilde k_{f_n(b)}(x)$ for all $x\in\Omega_n^{Gb}$ as shown before. Hence $\widetilde k_{f_{n-1}(a)}(\partial_n z)=\widetilde k_{f_n(b)}(z)$ for all $z\in C_n$. In particular $0=\widetilde k_{f_{n-1}(a)}(\partial_n\partial_{n+1}(y))=\widetilde k_{f_n(b)}(\partial_{n+1}(y))$ for all $y\in C_{n+1}$. Since $C_*$ is a $G$-complex, we also have $\widetilde k_{f_n(gb)}(\partial_{n+1}(y))=0$ for all $y\in C_{n+1}$ and all $g\in G$. Hence (2) is proven.

We obtain the following two $G$-subcomplexes ${\cal A}(Gb)$ and $C_*^{Gb}$:
\[\dots\longrightarrow 0\longrightarrow\span(Gb)\overset{\partial_n^{Gb}}\longrightarrow\span(Gf_{n-1}(a))\longrightarrow 0\longrightarrow\dots\]

\[\dots\longrightarrow C_{n+1}\overset{\partial_{n+1}}\longrightarrow\span(\Omega_n^{Gb})\overset{\partial_n}\longrightarrow\span(\Omega_{n-1}^{Gb})\overset{\partial_{n-1}}\longrightarrow C_{n-2}\longrightarrow\dots\]

Clearly we have $C_*\cong_G C_*^{Gb}\oplus{\cal A}(Gb)$. Furthermore $C_*^{Gb}$ is a free $G$-chain complex with $G$-basis $\Omega^{Gb}$, which is obtained from $\Omega$ by replacing $\Omega_n$ and $\Omega_{n-1}$ by $\Omega_n^{Gb}$ and $\Omega_{n-1}^{Gb}$ respectively.

Now we construct an $G$-equivariant acyclic matching $M^{Gb}$ on $P(C_*^{Gb},\Omega^{Gb})$ such that $w(b\succ a)$ is invertible for any $(a,b)\in M^{Gb}$. Let $R\subset P(C_*,\Omega)$ be the set of all elements appearing in any matching pair in $M\setminus G(a,b)$. We define the map
\begin{eqnarray*}
f:R&\longrightarrow&P(C_*^{Gb},\Omega^{Gb})\\
x&\longmapsto&f_n(x)\text{ for $x\in\Omega_n\setminus Gb$}\\
x&\longmapsto&x\text{ else}
\end{eqnarray*}
which has the following property: Let $x,y\in R$ such that $x\succ y$, then we have $w(f(x)\succ f(y))=w(x\succ y)$. This is clear for $x,y\not\in C_n$. It remains to prove the two other cases:

Case $x\in\Omega_n$: $\varphi(b)=q$ is minimal in $Q'$ by choice. We have $b\not\succ y$, since $b>y$ implies $\varphi(b)>\varphi(y)\in Q'$ which contradicts the minimality of $q$. Because $C_*^{Gb}$ is a $G$-complex, this implies $w(gb\succ y)=0$ for all $g\in G$. Hence $w(f(x)\succ f(y))=w(f(x)\succ y)=w(x\succ y)-\sum_{b'\in Gb} w(x\succ\partial_n b')\cdot w(b'\succ y)=w(x\succ y)$.

Case $y\in\Omega_n$: As mentioned (how $k_x$ behaves on $\widetilde\Omega$) in the proof of the injectivity of $f_n$, we have $\widetilde k_{f(y)}(\partial_{n+1} x)=k_y(\partial_{n+1} x)$. Hence $w(f(x)\succ f(y))=w(x\succ f(y))=w(x\succ y)$.

We set
\[
M^{Gb}:=\{(f(a'),f(b')\mid(a',b')\in M\setminus G(a,b)\}
\]
By the property of $f$, any cycle in $M^{Gb}$ would map to a cycle in $M$ under $f^{-1}$. It is easy to see that $M^{Gb}$ has all desired properties. By induction hypothesis there exists a $G$-chain complex $X_*$ such that
\begin{eqnarray*}
C_*^{Gb}&\cong_G&X_*\oplus\bigoplus_{G(a',b')\in M^{Gb}/G}{\cal A}(Gb')\\
&=&X_*\oplus\bigoplus_{G(a',b')\in M\setminus G(a,b)/G}{\cal A}(Gf_n(b'))
\end{eqnarray*}
$f_n$ induces a $G$-chain isomorphism $(\varphi_n)_n:{\cal A}(Gb')\longrightarrow{\cal A}(Gf_n(b'))$ for any $b'\in R$, where
\begin{eqnarray*}
\varphi_n:\span(Gb')&\longrightarrow&\span(Gf_n(b'))\\
x&\longmapsto&f_n(x)
\end{eqnarray*}
\begin{eqnarray*}
\varphi_{n-1}:\span(G\partial_nb')&\longrightarrow&\span(G\partial_nf_n(b'))\\
x&\longmapsto&(\partial_n\circ f_n)((\partial^{Gb'}_n)^{-1}(x))
\end{eqnarray*}
By Lemma \ref{iso}, the restrictions of $\partial_n$ used in the definition of $\varphi_{n-1}$ are isomorphisms.

We set $C_*^M:=X_*$.
\begin{eqnarray*}
C_*\cong_G C_*^{Gb}\oplus{\cal A}(Gb)&\cong_G&\left(C_*^M\oplus\bigoplus_{G(a',b')\in M\setminus G(a,b)/G}{\cal A}(Gb')\right)\oplus{\cal A}(Gb)\\
&=&C_*^M\oplus\bigoplus_{G(a',b')\in M/G}{\cal A}(Gb')
\end{eqnarray*}
\end{proof}
In the case $G$ is the trivial group, our result is the same as Theorem 11.24 in \cite[Chapter 11.3]{buch}, where $C_*^M$ is the Morse complex defined in \cite[Chapter 11.3]{buch}. By forgetting the action of $G$ we obtain
\[
\bigoplus_{G(a,b)\in M/G}{\cal A}(Gb)\cong\bigoplus_{(a,b)\in M}{\cal A}(b)
\]
Since the Morse complex for the matching $M$ is also a $G$-complex, it is usable in the $G$-equivariant case too.
\begin{cor}
$C_*$ is $G$-homotopy equivalent to $C_*^M$.
\end{cor}
\begin{proof}
$T_*$ is $G$-homotopy equivalent to the zero complex by Lemma \ref{iso}. Hence $C_*^M\oplus 0$ is $G$-homotopy equivalent to $C_*^M\oplus T_*$ which is $G$-isomorphic to $C_*$ by Proposition \ref{eqthm}, in particular $G$-homotopy equivalent.
\end{proof}
\bibliographystyle{model1-num-names}
\begin{thebibliography}{00}
\bibitem{donau3}
R. Donau, \textit{A nice acyclic matching on the nerve of the partition lattice}, \texttt{arXiv:1204.2693 [math.AT]} (2015).
\bibitem{freij}
R. Freij, \textit{Equivariant discrete Morse theory}, Discrete Mathematics \textbf{309} (12) (2009), pp. 3821-3829.
\bibitem{clmap}
D.N. Kozlov, \textit{Closure maps on regular trisps}, Topology and its Applications \textbf{156} (15) (2009), pp. 2491-2495.
\bibitem{buch}
D. Kozlov, \textit{Combinatorial Algebraic Topology}, Algorithms and Computation in Mathematics \textbf{21}, Springer-Verlag Berlin Heidelberg, 2008.
\end{thebibliography}
\end{document}

### 2018-05-17 paper1/paper.tex
mistakes in paper1 fixed

@@
 \end{thm}
-The proofs of Theorem \ref{patchwork} and \ref{morsemain} as well as further facts on Discrete Morse Theory can be found in \cite[Chapter 11]{buch}.
+The proofs of Theorems \ref{patchwork} and \ref{morsemain} as well as further facts on Discrete Morse Theory can be found in \cite[Chapter 11]{buch}.
 \section{The collapsible subtrisp}
@@
 We replace the block $A$ containing $1$ by the two blocks $\{1\}$ and $A\setminus\{1\}$ and take the class of this new partition as $\varphi(b)$.
-By applying Corollary 2.5 in \cite{clmap} and Proposition 7 in \cite{jule} it is easy to see that this map is well-defined and a closure map\footnote{see Definition 2.1 in \cite{clmap}}.
+By applying Corollary 2.5 in \cite{clmap} and Proposition 7 in \cite{jule} it is easy to see that this map is well-defined and a closure map\footnote{See Definition 2.1 in \cite{clmap}}.
 By Theorem 2.2 in \cite{clmap} we get an acyclic matching on ${\cal F}(\Delta^*)$ where a simplex $\sigma$ is critical if and only if  $V(\sigma)\subset R$.
@@
 Let $V_G$ be the set of all these special vertices.
-We define a Poset $P_G:=V_G\cup\{0\}$ such that $0$ is the smallest element of $P_G$ and the only element that is comparable with some other element.
+We define a poset $P_G:=V_G\cup\{0\}$ such that $0$ is the smallest element of $P_G$ and the only element that is comparable with some other element.
 That means $x,y\in P_G$, $x<y$ implies $x=0$.
@@
 Let $p=\{B_1,B_2,\dots,B_r\}$, where $B_1$ is the block that contains $1$, be a partition in the chain $\tilde\sigma$.
-For $p$ we define $p^*:=\{f_k(B_1)\cup\{k\},f_k(B_2),\dots,f_k(B_r)\}$, which is a partition of set set $[n]$.
+For $p$ we define $p^*:=\{f_k(B_1)\cup\{k\},f_k(B_2),\dots,f_k(B_r)\}$, which is a partition of the set $[n]$.
 $\tilde\sigma$ can be written as
@@
 
-By the discussion at the end of section \ref{allgemein} it remains to show that there exists an acyclic matching on $\varphi_G^{-1}(v)$ for each $v\in V_G$ such that all critical simplices have dimension $n-3$.
+By the discussion at the end of Section \ref{allgemein} it remains to show that there exists an acyclic matching on $\varphi_G^{-1}(v)$ for each $v\in V_G$ such that all critical simplices have dimension $n-3$.
 Let $v\in V_G$, then we choose a $2\leq k\leq n$ such that $v_k$ is a representative of $v$.
@@
 \[\C\cdot(S_1\times S_{n-1})=\bigoplus_{\sigma\in S_1\times S_{n-1}}\C\cdot\sigma\]
-are isomorphic as $(S_1\times S_{n-1})$-modules. The Transfer Theorem tells us
+are isomorphic as $(S_1\times S_{n-1})$-modules. The Transfer Theorem\footnote{See \cite{transfer}} tells us
 \[
@@
 The author would like to thank Dmitry N.
-Feichtner-Kozlov for this interesting problem and the support of this work, Juliane Lehmann for the helpful discussions and Giacomo d'Antonio for the hint that led to an improvement of section \ref{giacomo}.
+Feichtner-Kozlov for this interesting problem and the support of this work, Juliane Lehmann for the helpful discussions and Giacomo d'Antonio for the hint that led to an improvement of Section \ref{giacomo}.
 \bibliographystyle{model1-num-names}

### 2018-06-12 paper3/equivariant.tex
paper3v3 with just cosmetic changes

\documentclass{elsarticle}
\usepackage{amsmath,amsthm}
\usepackage{bbm}
\newtheorem{df}{Definition}[section]
\newtheorem{prop}[df]{Proposition}
\newtheorem{lem}[df]{Lemma}
\newtheorem{thm}[df]{Theorem}
\newtheorem{cor}[df]{Corollary}
\newtheorem{ex}[df]{Example}
\newtheorem{rem}[df]{Remark}
\def\N{\mathbbm N}
\def\Z{\mathbbm Z}
\def\id{\operatorname{id}}
%\def\card{\operatorname{card}}
%\def\C{\mathbbm C}
\journal{arXiv.org}
\begin{document}
\begin{frontmatter}
\title{A nice acyclic matching on the nerve of the partition lattice}
\author[ruelle]{Ralf Donau}
\address[ruelle]{Fachbereich Mathematik, Universit\"at Bremen, Bibliothekstra\ss e 1, 28359 Bremen, Germany}
\ead{ruelle@math.uni-bremen.de}
\begin{abstract}
The author has already proven that the space $\Delta(\Pi_n)/G$ is homotopy equivalent to a wedge of spheres of dimension $n-3$ for all natural numbers $n\geq 3$ and all subgroups $G\subset S_1\times S_{n-1}$. We wish to construct an acyclic matching on $\Delta(\Pi_n)/G$ that allows us to give a basis of its cohomology. This is also a more elementary approach to determining the number of spheres. Furthermore we give a description of the group action by an action on the spheres. We also obtain another result that we call Equivariant Patchwork Theorem.
\end{abstract}
\begin{keyword}
Discrete Morse Theory\sep Regular trisp\sep Acyclic matching\sep Equivariant homotopy
\end{keyword}
\end{frontmatter}
\section{Introduction}
Let $n\geq 3$ and let $\Pi_n$ denote the poset consisting of all partitions of $[n]:=\{1,\dots,n\}$ ordered by refinement, such that the finer partition is the smaller partition. Let $\overline{\Pi}_n$ denote the poset obtained from $\Pi_n$ by removing both the smallest and greatest element, which are $\{\{1\},\dots,\{n\}\}$ and $\{[n]\}$, respectively. We consider $\overline{\Pi}_n$ as a category, which is acyclic, and define $\Delta(\overline{\Pi}_n)$ to the nerve of the acyclic category $\overline{\Pi}_n$, which is a regular trisp, see \cite[Chapter 10]{buch}. The symmetric group $S_n$ operates on $\Delta(\overline{\Pi}_n)$ in a natural way.

It is well-known that $\Delta(\overline{\Pi}_n)$ is homotopy equivalent to a wedge of spheres of dimension $n-3$. The following two theorems are the first results concerning the topology of the quotient $\Delta(\overline{\Pi}_n)/G$, where $G$ is a non-trivial subgroup of $S_n$.
\begin{thm}[Kozlov, \cite{clmap}] % p. 198
For any $n\geq 3$, the topological space $\Delta(\overline{\Pi}_n)/S_n$ is contractible.
\end{thm}
We consider the Young subgroup $S_1\times S_{n-1}:=\{\sigma\in S_n\mid\sigma(1)=1\}$.
\begin{thm}[Donau, \cite{donau}]
\label{rdthm}
Let $n\geq3$ and $G\subset S_1\times S_{n-1}$ be a subgroup, then the topological space $\Delta(\overline{\Pi}_n)/G$ is homotopy equivalent to a wedge of $k$ spheres of dimension $n-3$, where $k$ is the index of $G$ in $S_1\times S_{n-1}$.
\end{thm}
This leads to a general question of determining the homotopy type of $\Delta(\overline{\Pi}_n)/G$ for an arbitrary subgroup $G\subset S_n$. One might conjecture that $\Delta(\overline{\Pi}_n)/G$ is homotopy equivalent to a wedge of spheres for any $n\geq 3$ and any subgroup $G\subset S_n$. But unfortunately this statement is not true as the following example will show: Let $p\geq 5$ be a prime number and let $C_p$ denote the subgroup of $S_p$ that is generated by the cycle $(1,2,\dots,p)$. Then the fundamental group of $\Delta(\overline{\Pi}_p)/C_p$ is isomorphic to $\Z/p\Z$. In particular $\Delta(\overline{\Pi}_p)/C_p$ cannot be homotopy equivalent to a wedge of spheres. A proof, which uses facts about covering spaces\footnote{See \cite[Chapter 1.3]{hatcher}}, can be found in \cite{torsion}.

%Another example are the Young subgroups of the form $G=S_k\times S_{n-k}$. Computer calculations of the homology reveal that $\Delta(\overline{\Pi}_n)/G$ is not homotopy equivalent to a wedge of spheres for $n=8$ and $G=S_2\times S_6$ or $G=S_4\times S_4$.

In this paper we construct an $(S_1\times S_{n-1})$-equivariant acyclic matching on the face poset ${\cal F}(\Delta(\overline{\Pi}_n))$ of $\Delta(\overline{\Pi}_n)$ for $n\geq 3$ such that we have a description of the critical simplices. This induces an acyclic matching on $\Delta(\overline{\Pi}_n)/G$ for any subgroup $G\subset S_1\times S_{n-1}$. Another benefit of having a description of the critical simplices is that we can easily give cocycle representatives of the generators of the cohomology, which can be useful for further analysis of the cohomology.

Equivariant acyclic matchings are also useful to find equivariant homotopies between spaces, since there exists an equivariant version of the Main Theorem of Discrete Morse Theory, see \cite{freij}. For the construction of an equivariant acyclic matching we have similar tools as in Discrete Morse Theory. An equivariant closure operator induces an equivariant trisp closure map which induces an equivariant acyclic matching. A detailed description of the non-equivariant versions of these tools can be found in \cite{clmap}.

Consider the Young subgroup $S_{k_1}\times S_{k_2}\times\dots\times S_{k_r}\subset S_n$ with $k_1+k_2+\dots+k_r=n$. During my work on this paper Gregory Arone\footnote{See Example 1.4 in \cite{arone}} discovered the following: $\Delta(\overline{\Pi}_n)$ is $(S_{k_1}\times S_{k_2}\times\dots\times S_{k_r})$-homotopy equivalent to a wedge of $(n-1)!$ spheres of dimension $n-3$ if $\gcd(k_1,k_2,\dots,k_r)=1$. $S_{k_1}\times S_{k_2}\times\dots\times S_{k_r}$ acts freely on these spheres. This is a generalization of Corollary \ref{moeglichkeit2} and Lemma \ref{freetrans}. Notice that a preprint of this paper was already published on arXiv in 2012.
\section{Discrete Morse Theory}
The definitions of regular trisps, partial matchings, acyclic matchings and foundations of Discrete Morse Theory can be found in \cite{forman,clmap,buch}. The following two theorems of Discrete Morse Theory are frequently used in our proofs.
\begin{thm}[Patchwork Theorem]
\label{patchwork}
Let $\varphi:P\longrightarrow Q$ be an order-preserving map and assume we have acyclic matchings on the subposets $\varphi^{-1}(q)$ for all $q\in Q$. Then the union of these matchings is an acyclic matching on $P$.
\end{thm}
\begin{thm}%[Main Theorem of Discrete Morse Theory for CW Complexes]
\label{morsemain}
Let $\Delta$ be a finite regular trisp and let $M$ be an acyclic matching on the poset ${\cal F}(\Delta)\setminus\{\hat{0}\}$. Let $c_i$ denote the number of critical $i$-dimensional simplices of $\Delta$. Then $\Delta$ is homotopy equivalent to a CW complex with $c_i$ cells of dimension $i$.
\end{thm}
The proofs of Theorems \ref{patchwork} and \ref{morsemain} as well as further facts on Discrete Morse Theory can be found in \cite[Chapter 11]{buch}.
\section{An Equivariant Patchwork Theorem}
We wish to construct an equivariant acyclic matching on a poset by gluing together smaller equivariant acyclic matchings on parts of the poset. This is similar to Theorem \ref{patchwork} with the difference that we also create copies of these matchings in our construction, see Figure \ref{eqex}.
\begin{df}
Let $P$ be a poset and let $G$ be a group acting on $P$. Let $M$ be an acyclic matching on $P$. We call $M$ an \emph{$G$-equivariant acyclic matching} if $(a,b)\in M$ implies $(ga,gb)\in M$ for all $g\in G$ and $a,b\in P$.
\end{df}
Let $G$ be a group acting on some posets $P$ and $Q$. For an element $q\in Q$ we set $G_q:=\{g\in G\mid gq=q\}$, known as the stabilizer subgroup of $q$.
\begin{prop}
\label{eqpatchwork}
Let $\varphi:P\longrightarrow Q$ be an order-preserving $G$-map and let $R\subset Q$ be a subset such that $R$ contains exactly one representative for each orbit in $Q$. Assume for each $r\in R$ we have an $G_r$-equivariant acyclic matching $M_r$ on $\varphi^{-1}(r)$. For $r\in R$, let $C_r$ denote the set of critical elements of $M_r$. Then we have an $G$-equivariant acyclic matching on $P$ such that
\[
\bigcup_{g\in G, r\in R}gC_r
\]
is the set of critical elements.

Let $r\in R$ and assume $G_r$ acts transitively on $C_r$. Then $G$ acts transitively on
\[
\bigcup_{g\in G}gC_r
\]
\end{prop}
\begin{proof}
We define acyclic matchings on the fibers of $\varphi$ as follows. For each $q\in Q$ we choose $r\in R$ and $g\in G$ with $gr=q$. The map $g:\varphi^{-1}(r)\longrightarrow\varphi^{-1}(q)$, which is an isomorphism of posets, induces an acyclic matching on $\varphi^{-1}(q)$. If we choose another $h\in G$ with $hr=q$, then we obtain the same matching. By Theorem \ref{patchwork} the union of these acyclic matchings is an acyclic matching which is $G$-equivariant by construction. The second statement is easy to see.
\end{proof}
\begin{figure}[ht]
\centering
\includegraphics{eqex.pdf}
\caption{A simple example: An $\Z_2$-equivariant acyclic matching composed of acyclic matchings on the fibers of $0$ and $1$. The matching pair in the fiber of $1$ is copied to the fiber of $2$. $\Z_2$ acts on both posets by reflection across the vertical line.}
\label{eqex}
\end{figure}
\begin{rem}
\label{quotmatching}
Let $G$ be a group acting on a finite regular trisp $\Delta$. Assume we have an $G$-equivariant acyclic matching $M$ on ${\cal F}(\Delta)\setminus\{\hat{0}\}$. Let $C$ be the set of critical simplices. Clearly we have an action of $G$ on $C$. Let $H\subset G$ be a subgroup. Then $M/H$ is an acyclic matching on ${\cal F}(\Delta/H)\setminus\{\hat{0}\}$, where $C/H$ is the set of critical simplices. In particular, if $\Delta$ is $G$-collapsible, then $\Delta/H$ is collapsible. Furthermore if $H$ is a normal subgroup, then the acyclic matching $M/H$ is $(G/H)$-equivariant.
\end{rem}
We also have an equivariant version of the Main Theorem of Discrete Morse Theory.
\begin{thm}[Freij, \cite{freij}]
\label{freij}
Let $G$ be a finite group. Let $\Delta$ be a finite regular $G$-trisp and let $M$ be a $G$-equivariant acyclic matching on the poset ${\cal F}(\Delta)\setminus\{\hat{0}\}$. Let $c_i$ denote the number of critical $i$-dimensional simplices of $\Delta$. Then $\Delta$ is $G$-homotopy equivalent to a $G$-CW complex where the cells correspond to the critical simplices of $M$ and the action of $G$ is the same as the action on $\Delta$ restricted to the critical simplices of $M$.
\end{thm}
There exists a characterization of acyclic matchings by means of order-preserving maps with small fibers, see Definition 11.3 and Theorem 11.4 in \cite[Chapter 11]{buch}. In a similar way we can also characterize $G$-equivariant acyclic matchings by means of order-preserving $G$-maps with small fibers.

For an order-preserving map with small fibers $\varphi$ let $M(\varphi)$ denote its associated acyclic matching which consists of all fibers of cardinality $2$, see \cite[Chapter 11]{buch}.
\begin{prop}
Let $G$ be a group acting on a finite poset $P$. For any order-preserving $G$-map $\varphi:P\longrightarrow Q$ with small fibers, the acyclic matching $M(\varphi)$ is $G$-equivariant. On the other hand, any $G$-equivariant acyclic matching $M$ on $P$ can be represented as $M=M(\varphi)$, where $\varphi:P\longrightarrow Q$ is an order-preserving $G$-map with small fibers.
\end{prop}
For two disjoint posets $(P_1,\leq_{P_1})$, $(P_2,\leq_{P_2})$ let $P_1\oplus P_2$ denote the linear sum of $P_1$ and $P_2$, where the set of $P_1\oplus P_2$ is $P_1\cup P_2$ and for two elements $x,y\in P_1\cup P_2$ we set $x\leq y$ if one of the following three conditions is satisfied:
\begin{enumerate}
\item $x,y\in P_1$ and $x\leq_{P_1} y$
\item $x,y\in P_2$ and $x\leq_{P_2} y$
\item $x\in P_1$ and $y\in P_2$
\end{enumerate}
Furthermore if a group $G$ acts on $P_1$ and on $P_2$, then we obtain a canonical action of $G$ on $P_1\oplus P_2$.
\begin{proof}
The first statement is easy to see. Assume we are given an $G$-equivariant acyclic matching $M$ on $P$.

We construct the poset $Q$ with an action of $G$ and the $G$-map $\varphi:P\longrightarrow Q$ inductively by adding either elements of $P$ or elements of $M$ to $Q$ in each step and extending the domain $D\subset P$ of the map $\varphi:D\longrightarrow Q$ such that $D$ has the following property: $d\geq p$ implies $p\in D$ for all $d\in D$ and all $p\in P$. We repeat these steps until we have reached $D=P$. We start with $Q=\emptyset$, $D=\emptyset$ and the trivial map $\varphi:D\longrightarrow Q$. Let $W$ denote the set of minimal elements in $P\setminus D$. We consider two cases.

First we assume one element $c\in W$ is critical. The orbit $Gc$ of $c$ is a subposet of $P$ where each element is only comparable to itself.
\begin{eqnarray*}
\widetilde\varphi:D\cup Gc&\longrightarrow&Q\oplus Gc\\
x&\longmapsto&
\begin{cases}
\varphi(x)&\text{for $x\in D$}\\
x&\text{for $x\in Gc$}\\
\end{cases}
\end{eqnarray*}
Assume $x,y\in D\cup Gc$ with $x\leq y$. $y\in D$ implies $x\in D$. $x\in D$ and $y\in Gc$ implies $\widetilde\varphi(x)\in Q$ and $\widetilde\varphi(y)\in Gc$. Hence $\widetilde\varphi$ is an order-preserving map. Now we assume $d\in D\cup Gc$, $p\in P$ with $d\geq p$. $d\in D$ in implies $p\in D$ by induction hypothesis. For $d\in Gc$ notice that $c$ is a minimal element in $P\setminus D$. Proceed with $Q=Q\oplus Gc$, $D=D\cup Gc$ and $\varphi=\widetilde\varphi$.

Now we assume that all elements in $W$ are matched. For $(a,b)\in M$ we write $a=d(b)$ and $b=u(a)$. We choose an $a\in W$ such that the only element in $W\cup u(W)$ that is smaller that $u(a)$ is $a$ itself. Such an element exists by the proof of Theorem 11.2 in \cite[Chapter 11]{buch}. We consider the orbit $G(a,u(a))$ of $(a,u(a))\in M$ as a poset where each element is only comparable to itself.
\begin{eqnarray*}
\widetilde\varphi:D\cup Ga\cup Gu(a)&\longrightarrow&Q\oplus G(a,u(a))\\
x&\longmapsto&
\begin{cases}
\varphi(x)&\text{for $x\in D$}\\
(x,u(x))&\text{for $x\in Ga$}\\
(d(x),x)&\text{for $x\in Gu(a)$}\\
\end{cases}
\end{eqnarray*}
We notice that $x<y$ implies $(x,y)\in M$ for all $x,y\in Ga\cup Gu(a)$ by the choice of $a$, hence $\widetilde\varphi(x)=\widetilde\varphi(y)$. $\widetilde\varphi$ is order-preserving and $D$ has the desired property by almost the same arguments as in the first case. Proceed with $Q=Q\oplus G(a,u(a))$, $D=D\cup Ga\cup Gu(a)$ and $\varphi=\widetilde\varphi$.

$\varphi$ is an order-preserving $G$-map with small fibers by construction. Clearly we have $M=M(\varphi)$. Notice that $\varphi$ is even surjective.
\end{proof}
\section{The main result}
Let $n\geq 3$ be a fixed natural number.
\begin{df}
Let $A$ be the set of all vertices of $\Delta(\overline{\Pi}_n)$ where all blocks not containing $1$ are singleton. We define the following set of simplices of $\Delta(\overline{\Pi}_n)$, see Figure \ref{exsimplex}.
\[
C_n:=\{\sigma\in{\cal F}(\Delta(\overline{\Pi}_n))\mid\text{$V(\sigma)\subset A$ and $\dim\sigma=n-3$}\}
\]
$V(\sigma)$ denotes the set of vertices of $\sigma$ and ${\cal F}(\Delta(\overline{\Pi}_n))$ denotes the face poset of $\Delta(\overline{\Pi}_n)$. Furthermore we set $\alpha_n$ to the vertex $\{\{1\},\{2,\dots,n\}\}$.
\end{df}
\begin{figure}[ht]
\centering
\includegraphics{simplex.pdf}
\caption{A simplex in $C_5$ which has dimension $2$.}
\label{exsimplex}
\end{figure}
\begin{rem}
\label{card}
The cardinality of $C_n$ is $(n-1)!$.
\end{rem}
\begin{prop}
\label{main}
There exists an $(S_1\times S_{n-1})$-equivariant acyclic matching on ${\cal F}(\Delta(\overline{\Pi}_n))$, such that $C_n\cup\{\alpha_n\}$ is the set of critical simplices.
\end{prop}
Let $V$ be the set of all vertices where the block containing $1$ has exactly two elements and any other block is singleton. Such a vertex can be written as
\[
v_k:=\{\{1,k\},\{2\},\dots,\widehat{\{k\}},\dots,\{n\}\}
\]
with $k\in\{2,\ldots, n\}$. The element with the hat above is omitted.

We define a poset $P:=V\cup\{0\}$ such that $0$ is the smallest element of $P$ and the only element that is comparable with some other element. That means $x,y\in P$, $x<y$ implies $x=0$. We define an order-preserving map $\varphi:{\cal F}(\Delta(\overline{\Pi}_n))\longrightarrow P$ as follows. Let $\sigma\in{\cal F}(\Delta(\overline{\Pi}_n))$, then we map $\sigma$ to $0$ if $V(\sigma)\cap V=\emptyset$. Otherwise we map $\sigma$ to the special vertex of $V$ that belongs to $\sigma$, which is unique. Notice that $S_1\times S_{n-1}$ acts on $P$ in a natural way and $\varphi$ is a $(S_1\times S_{n-1})$-map. $P$ has two orbits where one consists of one element which is $0$. The other orbit may be represented by $v_n$.
\begin{lem}
\label{mitte}
There exists an $(S_1\times S_{n-1})$-equivariant acyclic matching on $\varphi^{-1}(0)$, such that $\alpha_n$ is the only critical simplex.
\end{lem}
\begin{proof}
The proof of Lemma \ref{mitte} is the same as the proof of Lemma 3.2 in \cite{donau} for the case $G=\{\id_{[n]}\}$. It is easy to see that the acyclic matching, that is constructed in this proof, is $(S_1\times S_{n-1})$-equivariant.
\end{proof}
\begin{proof}[Proof of Proposition \ref{main}]
It is easy to see, that the statement is true for $n=3$. Now we assume $n>3$ and proceed by induction.

We define acyclic matchings on $\varphi^{-1}(0)$ and $\varphi^{-1}(v_n)$ as follows. By Lemma \ref{mitte} there exists an $(S_1\times S_{n-1})$-equivariant acyclic matching on $\varphi^{-1}(0)$, where $\alpha_n$ is the only critical simplex.

We define a map
\[
\psi:{\cal F}(\Delta(\overline{\Pi}_{n-1}))\longrightarrow\varphi^{-1}(v_n)\setminus\{v_n\}
\]
as follows. We add $n$ to the block that contains $1$ in each partition and append $v_n$ to the bottom of the chain, see Figure \ref{exmapping}. The map $\psi$ is an isomorphism of posets. A similar definition of $\psi$ as well as a detailed description of its inverse can be found in the proof of Lemma 4.1 in \cite{donau}.
\begin{figure}[ht]
\centering
\includegraphics{mapping.pdf}
\caption{Example for the map $\psi$, where $n=5$.} % ERKLAERUNG ZUM BILD FEHLT
\label{exmapping}
\end{figure}

Via $\psi$ we get an acyclic matching $M$ on $\varphi^{-1}(v_n)$, where the set of critical simplices consists of the simplices in $\psi[C_{n-1}]$, one critical simplex $s_n$ consisting of the two vertices $v_n$ and $\{\{1,n\},\{2,\dots,(n-1)\}\}$, which has dimension $1$. Additionally we have the critical simplex that has only the vertex $v_n$. Finally we match $v_n$ with $s_n$.

We have to show that $\sigma(v_n)=v_n$ and $(a,b)\in M$ implies $(\sigma a,\sigma b)\in M$ for all $\sigma\in S_1\times S_{n-1}$ and all $a,b\in\varphi^{-1}(v_n)$. Let $\sigma\in S_1\times S_{n-1}$. $\sigma(v_n)=v_n$ implies $\sigma(1)=1$ and $\sigma(n)=n$. We define a $\widetilde\sigma\in S_1\times S_{n-2}$ by setting $\widetilde\sigma(x):=\sigma(x)$ for $1\leq x\leq n-1$. Notice $\sigma\psi=\psi\widetilde\sigma$ which implies $\psi^{-1}\sigma=\widetilde\sigma\psi^{-1}$. Let $(a,b)\in M$. Clearly we have $(v_n,s_n)=(\sigma v_n,\sigma s_n)$. Now we assume $a\not=v_n$ and $b\not=s_n$. By induction hypothesis, we have an acyclic matching $\widetilde M$ on ${\cal F}(\Delta(\overline{\Pi}_{n-1}))$ which is $(S_1\times S_{n-2})$-equivariant. By the construction of $M$ we have $(\psi(a)^{-1},\psi(b)^{-1})\in\widetilde M$. This implies $(\widetilde\sigma\psi(a)^{-1},\widetilde\sigma\psi(b)^{-1})\in\widetilde M$, hence $(\sigma a,\sigma b)\in M$.

By Proposition \ref{eqpatchwork} there exists an $(S_1\times S_{n-1})$-equivariant acyclic matching on ${\cal F}(\Delta(\overline{\Pi}_n))$ such that
\[
\left(\bigcup_{g\in S_1\times S_{n-1}}g\psi[C_{n-1}]\right)\cup\{\alpha_n\}
\]
is the set of critical elements. It is easy to see that this set equals $C_n\cup\{\alpha_n\}$.
\end{proof}
\begin{cor}
\label{moeglichkeit1}
Let $G\subset S_1\times S_{n-1}$ be a subgroup. Then there exists an acyclic matching on ${\cal F}(\Delta(\overline{\Pi}_n))/G$, such that the set of critical simplices consists of the simplices in $C_n/G$ and $\alpha_n/G$.
\end{cor}
\begin{proof}
Apply Remark \ref{quotmatching}.
\end{proof}
\begin{ex}
\label{nicematchingex}
Assume $G=S_1\times S_{n-1}$. The vertices of $\Delta(\overline{\Pi}_n)/S_1\times S_{n-1}$ can be indexed with number partitions of $n$, which we may write as $v_0\oplus v_1+\dots+v_r$, that distinguish the first number, i.e.\ $\oplus$ is non-commutative. The number on the left side of $\oplus$, that is $v_0$, corresponds to the block that contains $1$. There exists an acyclic matching on the poset ${\cal F}(\Delta(\overline{\Pi}_n)/S_1\times S_{n-1})$, where the set of critical simplices consists of the vertex $1\oplus(n-1)$ and the unique simplex $\sigma$ whose vertices are $v_0\oplus 1^{n-v_0}$ with $v_0=2,\dots,n-1$, which has dimension $n-3$.
\end{ex}
A slightly different proof of the result in Example \ref{nicematchingex}, as well as a detailed description of $\Delta(\overline{\Pi}_n)/S_1\times S_{n-1}$, can be found in \cite{donau2}.
\section{Applications}
Let $n\geq 3$ be a fixed natural number.
\begin{cor}
\label{moeglichkeit2}
The topological space $\Delta(\overline{\Pi}_n)$ is $(S_1\times S_{n-1})$-homotopy equivalent to a wedge of $(n-1)!$ spheres of dimension $n-3$. The spheres are indexed with the simplices in $C_n$, which induces an action of $S_1\times S_{n-1}$ on the $(n-1)!$ spheres.
\end{cor}
\begin{proof}
Apply Theorem \ref{freij}.
\end{proof}
\begin{lem}
\label{freetrans}
$S_1\times S_{n-1}$ acts freely and transitively on $C_n$.
\end{lem}
\begin{proof}
Since $C_n=\bigcup_{g\in S_1\times S_{n-1}}g\psi[C_{n-1}]$, the action is transitive, which follows inductively by the second statement of Proposition \ref{eqpatchwork}.

By Remark \ref{card} the cardinality of $C_n$ is $(n-1)!$ which equals the cardinality of $S_1\times S_{n-1}$. Hence the action is free.
\end{proof}
Let $G\subset S_1\times S_{n-1}$ be an arbitrary subgroup.
\begin{rem}
The cardinality of $C_n/G$ is the index of $G$ in $S_1\times S_{n-1}$.
\end{rem}
\begin{proof}
Apply Lemma \ref{freetrans}.
\end{proof}
Now Theorem \ref{rdthm} follows as a corollary. We can either apply Corollary \ref{moeglichkeit1} or Corollary \ref{moeglichkeit2}.
\begin{cor}
The topological space $\Delta(\overline{\Pi}_n)/G$ is homotopy equivalent to a wedge of spheres of dimension $n-3$. The number of spheres is the index of $G$ in $S_1\times S_{n-1}$.
\end{cor}
\section*{Acknowledgments}
The author would like to thank Dmitry N. Kozlov for this interesting problem, Ragnar Freij and Giacomo d'Antonio for the helpful discussions.
\bibliographystyle{model1-num-names}
\begin{thebibliography}{00}
\bibitem{arone}
G. Arone, \textit{A branching rule for partition complexes}, \texttt{arXiv:1508.05086v1 [math.AT]} (2015).
\bibitem{donau2}
R. Donau, \textit{On a quotient topology of the partition lattice with forbidden block sizes}, Topology and its Applications \textbf{159} (8) (2012), pp. 2052-2057.
\bibitem{donau}
R. Donau, \textit{Quotients of the order complex $\Delta(\overline{\Pi}_n)$ by subgroups of the Young subgroup $S_1\times S_{n-1}$}, Topology and its Applications \textbf{157} (16) (2010), pp. 2476-2479.
\bibitem{torsion}
R. Donau, \textit{Quotients of the topology of the partition lattice which are not homotopy equivalent to wedges of spheres}, \texttt{arXiv:1202.4368v2 [math.AT]} (2012).
\bibitem{forman}
R. Forman, \textit{Morse theory for cell complexes}, Adv. Math. \textbf{134} (1) (1998), pp. 90-145.
\bibitem{freij}
R. Freij, \textit{Equivariant discrete Morse theory}, Discrete Mathematics \textbf{309} (12) (2009), pp. 3821-3829.
\bibitem{hatcher}
A. Hatcher, \textit{Algebraic Topology}, Cambridge University Press, 2008.
\bibitem{clmap}
D.N. Kozlov, \textit{Closure maps on regular trisps}, Topology and its Applications \textbf{156} (15) (2009), pp. 2491-2495.
\bibitem{buch}
D. Kozlov, \textit{Combinatorial Algebraic Topology}, Algorithms and Computation in Mathematics \textbf{21}, Springer-Verlag Berlin Heidelberg, 2008.
\end{thebibliography}
\end{document}

### 2018-12-17 paper4/eqalg.tex
paper4 improved

@@
 \[\dots\longrightarrow 0\longrightarrow\span(Gb)\overset{\partial^{Gb}_n}\longrightarrow\span(G\partial_n b)\longrightarrow 0\longrightarrow\dots\]
-$\partial^{Gb}_n$ denotes the restriction of $\partial_n$ to $\span(Gb)$, i.e. $\partial^{Gb}_n(x):=\partial_n(x)$.
+$\partial^{Gb}_n$ denotes the restriction of $\partial_n:C_n\longrightarrow C_{n-1}$ to $\span(Gb)$, i.e. $\partial^{Gb}_n(x):=\partial_n(x)$ for $x\in\span(Gb)$.
 \end{df}
@@
 
-The composition ${\cal A}(Gb)\longrightarrow0\longrightarrow{\cal A}(Gb)$ is homotop to $\id:{\cal A}(Gb)\longrightarrow{\cal A}(Gb)$ via the $G$-chain homotopy
-\begin{eqnarray*}
-P:{\cal A}(Gb)&\longrightarrow&{\cal A}(Gb)\\
-x&\longmapsto&(\partial^{Gb}_n)^{-1}(x)\text{ for $x\in\span(G\partial_n b)$}\\
-x&\longmapsto&0\text{ else}
-\end{eqnarray*}
+The composition ${\cal A}(Gb)\longrightarrow0\longrightarrow{\cal A}(Gb)$ is homotop to $\id:{\cal A}(Gb)\longrightarrow{\cal A}(Gb)$ via the $G$-chain homotopy $P=(P_i)$, where
+\begin{itemize}
+\item $P_{n-1}:=(\partial^{Gb}_n)^{-1}$
+\item $P_i:\equiv 0$ for $i\not=n-1$
+\end{itemize}
 On the other hand $0\longrightarrow{\cal A}(Gb)\longrightarrow 0$ equals $\id:0\longrightarrow 0$. Hence ${\cal A}(Gb)$ is $G$-homotopy equivalent to the zero complex.
@@
 \[\dots\longrightarrow 0\longrightarrow\span(Gb)\overset{\partial_n^{Gb}}\longrightarrow\span(Gf_{n-1}(a))\longrightarrow 0\longrightarrow\dots\]
-
 \[\dots\longrightarrow C_{n+1}\overset{\partial_{n+1}}\longrightarrow\span(\Omega_n^{Gb})\overset{\partial_n}\longrightarrow\span(\Omega_{n-1}^{Gb})\overset{\partial_{n-1}}\longrightarrow C_{n-2}\longrightarrow\dots\]

### 2018-12-22 paper3/equivariant.tex
refs in paper3

@@
 \bibitem{torsion}
-R. Donau, \textit{Quotients of the topology of the partition lattice which are not homotopy equivalent to wedges of spheres}, \texttt{arXiv:1202.4368v2 [math.AT]} (2012).
+R. Donau, \textit{Quotients of the topology of the partition lattice which are not homotopy equivalent to wedges of spheres}, \texttt{arXiv:1202.4368 [math.AT]} (2012).
 \bibitem{forman}
@@
 D. Kozlov, \textit{Combinatorial Algebraic Topology}, Algorithms and Computation in Mathematics \textbf{21}, Springer-Verlag Berlin Heidelberg, 2008.
+\bibitem{wachs}
+M.L. Wachs, \textit{On the (co)homology of the partition lattice and the free Lie algebra}, Discrete Mathematics \textbf{193} (1-3) (1998), pp. 287-319.
 \end{thebibliography}

### 2019-01-31 paper3/equivariant.tex
linear sum -> ordinal sum

@@
 \end{prop}
-For two disjoint posets $(P_1,\leq_{P_1})$, $(P_2,\leq_{P_2})$ let $P_1\oplus P_2$ denote the linear sum of $P_1$ and $P_2$, where the set of $P_1\oplus P_2$ is $P_1\cup P_2$ and for two elements $x,y\in P_1\cup P_2$ we set $x\leq y$ if one of the following three conditions is satisfied:
+For two disjoint posets $(P_1,\leq_{P_1})$, $(P_2,\leq_{P_2})$ let $P_1\oplus P_2$ denote the \emph{ordinal sum}\footnote{Called \emph{stack} in \cite{buch}} of $P_1$ and $P_2$, where the set of $P_1\oplus P_2$ is $P_1\cup P_2$ and for two elements $x,y\in P_1\cup P_2$ we set $x\leq y$ if one of the following three conditions is satisfied:
 \begin{enumerate}

### 2019-05-16 paper2/paper.tex
mistakes in paper2 fixed
extract for all, mistake in paper2

@@
 \begin{df}
-A map $\varphi:B\longrightarrow R$ is called a \emph{closure map} if for any $\sigma\in{\cal F}(\Delta)\setminus{\cal F}(\Delta^R)$ either $m(\sigma)\in V(\sigma)$ or $m(\sigma)$ can be uniquely inserted into $\sigma$, i.e.
-there exists a unique simplex $\tau\in{\cal F}(\Delta)$ with $m(\sigma)\in V(\tau)$ and $\sigma$ is a subsimplex of $\tau$ with $\dim\tau=\dim\sigma+1$.
+A map $\varphi:B\longrightarrow R$ is called a \emph{closure map} if for any $\sigma\in{\cal F}(\Delta)\setminus{\cal F}(\Delta^R)$ either $\varphi(m(\sigma))\in V(\sigma)$ or $\varphi(m(\sigma))$ can be uniquely inserted into $\sigma$, i.e.
+there exists a unique simplex $\tau\in{\cal F}(\Delta)$ with $\varphi(m(\sigma))\in V(\tau)$ and $\sigma$ is a subsimplex of $\tau$ with $\dim\tau=\dim\sigma+1$.
 \end{df}
@@
 \input{forest.pdf_t} 
-\caption{A simplex of $\Delta(\Pi_{6,3})/S_1\times S_5$. One number is distinguished on each level.}
+\caption{A simplex of $\Delta(\Pi_{6,2})/S_1\times S_5$. One number is distinguished on each level.}
 \end{figure}
@@
 \centering
-\input{mapping.pdf_t} 
+\input{mapping_paper2.pdf_t}
 \caption{Example for the trisp isomorphism}
@@
 \]
-We give is a short description of $\partial^*$. Let $\tau$ be simplex of dimension $l+1$.
+We give a short description of $\partial^*$. Let $\tau$ be simplex of dimension $l+1$.
 \[

### 2019-07-15 paper3/equivariant.tex
extract for all, mistake in paper2
k, n-k

@@
 It is well-known that $\Delta(\overline{\Pi}_n)$ is homotopy equivalent to a wedge of spheres of dimension $n-3$.
-The following two theorems are the first results concerning the topology of the quotient $\Delta(\overline{\Pi}_n)/G$, where $G$ is a non-trivial subgroup of $S_n$.
+The following two theorems are results concerning the topology of the quotient $\Delta(\overline{\Pi}_n)/G$, where $G$ is a non-trivial subgroup of $S_n$.
 \begin{thm}[Kozlov, \cite{clmap}] % p. 198
@@
 
-Consider the Young subgroup $S_{k_1}\times S_{k_2}\times\dots\times S_{k_r}\subset S_n$ with $k_1+k_2+\dots+k_r=n$.
-During my work on this paper Gregory Arone\footnote{See Example 1.4 in \cite{arone}} discovered the following: $\Delta(\overline{\Pi}_n)$ is $(S_{k_1}\times S_{k_2}\times\dots\times S_{k_r})$-homotopy equivalent to a wedge of $(n-1)!$ spheres of dimension $n-3$ if $\gcd(k_1,k_2,\dots,k_r)=1$.
-$S_{k_1}\times S_{k_2}\times\dots\times S_{k_r}$ acts freely on these spheres.
+Consider the Young subgroup $S_k\times S_{n-k}\subset S_n$ with $1\leq k<n$.
+During my work on this paper Gregory Arone\footnote{See Example 1.4 in \cite{arone}} discovered the following: $\Delta(\overline{\Pi}_n)$ is $(S_k\times S_{n-k})$-homotopy equivalent to a wedge of $(n-1)!$ spheres of dimension $n-3$ if $\gcd(k,n-k)=1$.
+$S_k\times S_{n-k}$ acts freely on these spheres.
 This is a generalization of Corollary \ref{moeglichkeit2} and Lemma \ref{freetrans}.
-Notice that a preprint of this paper was already published on arXiv in 2012.
+Notice that the first preprint of this paper was already published on arXiv in 2012.
 \section{Discrete Morse Theory}
@@
 \begin{prop}
+\label{smallfibers}
 Let $G$ be a group acting on a finite poset $P$.
@@
 \bibitem{arone}
-G. Arone, \textit{A branching rule for partition complexes}, \texttt{arXiv:1508.05086v1 [math.AT]} (2015).
+G. Arone, \textit{A branching rule for partition complexes}, \texttt{arXiv:1508.05086 [math.AT]} (2015).
 \bibitem{donau2}

### 2019-11-21 paper3/equivariant.tex
Gerrit fix

@@
 The number on the left side of $\oplus$, that is $v_0$, corresponds to the block that contains $1$.
-There exists an acyclic matching on the poset ${\cal F}(\Delta(\overline{\Pi}_n)/S_1\times S_{n-1})$, where the set of critical simplices consists of the vertex $1\oplus(n-1)$ and the unique simplex $\sigma$ whose vertices are $v_0\oplus 1^{n-v_0}$ with $v_0=2,\dots,n-1$, which has dimension $n-3$.
+There exists an acyclic matching on the poset ${\cal F}(\Delta(\overline{\Pi}_n)/S_1\times S_{n-1})$, where the set of critical simplices consists of the vertex $1\oplus(n-1)$ and the simplex $\sigma$ of dimension $n-3$ whose vertices are $v_0\oplus 1^{n-v_0}$ with $v_0=2,\dots,n-1$.
 \end{ex}

### 2020-02-14 paper2/paper.tex
paper1 refto fix
paper2 cosmetic fix

@@
 \end{eqnarray*}
-By Lemma 3.2 in \cite{donau} we already have an acyclic matching on $\varphi^{-1}(0)$ with one critical simplex which has dimension $0$.
+By Lemma 3.2 in \cite{donau}
+we already have an acyclic matching on $\varphi^{-1}(0)$ with one critical simplex which has dimension $0$.
 The next step is to find an acyclic matching on $\varphi^{-1}(1)$ and apply the Patchwork Theorem, see \cite[Chapter 11]{buch}.
@@
 We define a map $\psi:{\cal F}(\Delta(\Pi_{n-1,2})/S_1\times S_{n-2})\longrightarrow\varphi^{-1}(1)\setminus\{2\oplus 1^{n-2}\}$, which increases the number on the left side by $1$ on each level and appends the vertex $2\oplus 1^{n-2}$ at the bottom of the forest.
-The map $\psi$ is an isomorphism between posets, see Lemma 4.1 in \cite{donau}, and $\psi$ induces an acyclic matching on $\varphi^{-1}(1)$ where the set of critical simplices consists of one simplex $\alpha$ of dimension $1$, the simplex that has only the vertex $2\oplus 1^{n-2}$ and the unique simplex $\sigma_n$ with the vertices $v_0\oplus 1^{n-v_0}$ with $v_0=2,\dots,n$ by induction hypothesis.
+The map $\psi$ is an isomorphism between posets,
+see Lemma 4.1 in \cite{donau},
+and $\psi$ induces an acyclic matching on $\varphi^{-1}(1)$ where the set of critical simplices consists of one simplex $\alpha$ of dimension $1$, the simplex that has only the vertex $2\oplus 1^{n-2}$ and the unique simplex $\sigma_n$ with the vertices $v_0\oplus 1^{n-v_0}$ with $v_0=2,\dots,n$ by induction hypothesis.
 Finally we match $2\oplus 1^{n-2}$ with $\alpha$.
@@
 The homomorphism $\rho_{n,k}$ is not a coboundary by the following argument: Assume there exists $g\in C^{l-1}$ with $\rho_{n,k}=\partial^*g$.
-If we restrict $\rho_{n,k}$ to $\Delta_{n,k}$, this implies $\sigma_{n_k}=\partial^*\widetilde g$, where $\widetilde g$ is the restriction of $g$ to $\Delta_{n,k}$.
+If we restrict $\rho_{n,k}$ to $\Delta_{n,k}$, this implies $\sigma_{n,k}=\partial^*\widetilde g$, where $\widetilde g$ is the restriction of $g$ to $\Delta_{n,k}$.
 It remains to show that $\rho_{n,k}$ is a generator.
@@
 \end{proof}
-%\section*{Acknowledgements}
-%The author would like to thank ...
 \bibliographystyle{model1-num-names}

### 2020-02-15 paper3/equivariant.tex
paper3 citation in theorems
paper3 splitting basis

@@
 \def\Z{\mathbbm Z}
+\def\C{\mathbbm C}
 \def\id{\operatorname{id}}
 %\def\card{\operatorname{card}}
-%\def\C{\mathbbm C}
 \journal{arXiv.org}
@@
 The following two theorems are results concerning the topology of the quotient $\Delta(\overline{\Pi}_n)/G$, where $G$ is a non-trivial subgroup of $S_n$.
-\begin{thm}[Kozlov, \cite{clmap}] % p. 198
+\begin{thm}[Kozlov \cite{clmap}]
 For any $n\geq 3$, the topological space $\Delta(\overline{\Pi}_n)/S_n$ is contractible.
@@
 We consider the Young subgroup $S_1\times S_{n-1}:=\{\sigma\in S_n\mid\sigma(1)=1\}$.
-\begin{thm}[Donau, \cite{donau}]
+\begin{thm}[Donau \cite{donau}]
 \label{rdthm}
@@
 We also have an equivariant version of the Main Theorem of Discrete Morse Theory.
-\begin{thm}[Freij, \cite{freij}]
+\begin{thm}[Freij \cite{freij}]
 \label{freij}
@@
 \end{cor}
+Now we consider the top cohomology $H^{n-3}(\Delta(\overline{\Pi}_n);\C)$ of $\Delta(\overline{\Pi}_n)$.
+For each simplex $\sigma\in C_n$ we obtain a $\sigma^*\in H^{n-3}(\Delta(\overline{\Pi}_n);\C)$ which is represented by a cocycle that maps $\sigma$ to $1$ and all other simplices to $0$.
+We obtain a basis $\{\sigma^*\mid\sigma\in C_n\}$ of the top cohomology which is called \emph{splitting basis} in \cite{wachs}.
 \section*{Acknowledgments}

### 2020-08-27 figures/torus3d.sh
remove alpha

@@ -1,3 +1,3 @@
 #!/bin/sh
 
-povray torus3d.pov +W1200 +H900 +A +UA
+povray torus3d.pov +W1200 +H900 +A

### 2020-08-28 paper2/paper.tex
paper2 small fix

@@
 The homomorphism $\rho_{n,k}$ is not a coboundary by the following argument: Assume there exists $g\in C^{l-1}$ with $\rho_{n,k}=\partial^*g$.
-If we restrict $\rho_{n,k}$ to $\Delta_{n,k}$, this implies $\sigma_{n,k}=\partial^*\widetilde g$, where $\widetilde g$ is the restriction of $g$ to $\Delta_{n,k}$.
+If we restrict $\rho_{n,k}$ to $\Delta_{n,k}$, this implies $\delta_{\sigma_{n,k}}=\partial^*\widetilde g$, where $\widetilde g$ is the restriction of $g$ to $\Delta_{n,k}$.
 It remains to show that $\rho_{n,k}$ is a generator.
 Assume there exists $f\in\hom(C_l,\Z)$ such that $[\rho_{n,k}]=k[f]$ with $k>1$.
-Again we restrict to $\Delta_{n,k}$ and obtain $[\sigma_{n,k}]=k[\widetilde f]$.
+Again we restrict to $\Delta_{n,k}$ and obtain $[\delta_{\sigma_{n,k}}]=k[\widetilde f]$.
 By the discussion above this is a contradiction.

### 2020-09-29 paper3/equivariant.tex
paper3v4

@@
 The author has already proven that the space $\Delta(\Pi_n)/G$ is homotopy equivalent to a wedge of spheres of dimension $n-3$ for all natural numbers $n\geq 3$ and all subgroups $G\subset S_1\times S_{n-1}$.
-We wish to construct an acyclic matching on $\Delta(\Pi_n)/G$ that allows us to give a basis of its cohomology.
+We construct an $S_1\times S_{n-1}$-equivariant acyclic matching on $\Delta(\Pi_n)$ together with a description of its critical simplices.
 This is also a more elementary approach to determining the number of spheres.
-Furthermore we give a description of the group action by an action on the spheres.
-We also obtain another result that we call Equivariant Patchwork Theorem.
+We also develop new methods for Equivariant Discrete Morse Theory by adapting the Patchwork Theorem and poset maps with small fibers from Discrete Morse Theory.
 \end{abstract}
@@
 This induces an acyclic matching on $\Delta(\overline{\Pi}_n)/G$ for any subgroup $G\subset S_1\times S_{n-1}$.
-Another benefit of having a description of the critical simplices is that we can easily give cocycle representatives of the generators of the cohomology, which can be useful for further analysis of the cohomology.
 
@@
 \section*{Acknowledgments}
-The author would like to thank Dmitry N. Kozlov for this interesting problem, Ragnar Freij and Giacomo d'Antonio for the helpful discussions.
+The author would like to thank Dmitry N.
+Kozlov for this interesting problem, Ragnar Freij and Giacomo d'Antonio for the helpful discussions.
+Furthermore the author would like to thank another anonymous reader of a previous version of this paper.
 \bibliographystyle{model1-num-names}
